---
title: "Houston Solar Panels 2022 version"
author: "Alan Jackson"
date: '2022-05-10'
slug: [Solar-panels-2022]
categories: 
- Permits
tags:
- Houston
- Mapping
keywords: tech
---

```{r setup, include=FALSE}

library(tidyverse)
library(ggplot2)
library(lubridate)
library(leaflet)

googlecrs <- 4326

path <- "/home/ajackson/Dropbox/Rprojects/Curated_Data_Files/Houston_Permits/"

df <- readRDS(paste0(path, "Clean_Final_2017_to_Mar_2022.rds"))
old <- readRDS(paste0(path, "Clean_Final_1989_2019.rds"))

knitr::opts_chunk$set(echo = TRUE)
```

Having bought solar panels myself a couple of years ago, and realizing
that the city permit database could be used to find most installations,
I decided that it would be interesting to look at the recent history and
a few other facets of residential solar panel installations.

The first step is to download the structural permit data as a CSV file
from [the city open data
website.](https://cohgis-mycity.opendata.arcgis.com/datasets/coh-permits-structural-bldg).
This file is no longer available, so I now download the data from the
[new
site](https://www.houstontx.gov/planning/DevelopRegs/docs_pdfs/Permit_eReport/)
and clean it up. Details may be found on my [github
site](https://github.com/alankjackson/Curated_Data/blob/main/Clean_City_Permit_Addresses.Rmd).
The code for cleaning up the older data is in the file
Clean_old_City_permits.Rmd on that same github site.

## Grabbing the correct records

As far as I can tell, Solar Panels are designated as such in the
Description field, and nothing else. So a simple filter on "Solar"
should suffice to capture all the installation permits. In more recent
records, batteries are also flagged with tags like BATTERY, BATTERIES,
LITHIUM, LI-ION, or POWERWALL. Additionally, since about 2017 but more
consistently in recent years, the number of panels has been noted, so I
can use that as a way to estimate added capacity.

And there are 2 installations of solar roof tiles.

```{r grab solar, message=FALSE, warning=FALSE, echo=FALSE}

solar <- df %>% 
  filter(stringr::str_detect(Comments, "SOLAR ")) %>% 
  filter(Permit_date>"2013-12-31") %>% 
  select(Address=Cleaned_address, Zipcode, Permit_type, Permit_date,
         Comments, Lon, Lat)

foo <- old %>% 
  filter(stringr::str_detect(Comments, "SOLAR ")) %>% 
  filter(Date>"2013-12-31") %>% 
  mutate(Zipcode=as.character(Zipcode)) %>% 
  mutate(Permit_type="") %>% 
  mutate(Date=as_date(Date)) %>% 
  select(Address=Output, Zipcode, Permit_type, Permit_date=Date,
         Comments, Lon, Lat) %>% 
  filter(Permit_date<"2017-01-01")

solar <- rbind(solar, foo) %>% as_tibble() %>% unique()

generator <- df %>% 
  filter(stringr::str_detect(Comments, "GAS GEN|DIESEL")) %>% 
  filter(!stringr::str_detect(Comments, "CELL")) %>% 
  filter(Permit_date>"2013-12-31") %>% 
  select(Address=Cleaned_address, Zipcode, Permit_type, Permit_date,
         Comments, Lon, Lat)

foo <- old %>% 
  filter(stringr::str_detect(Comments, "GAS GEN|DIESEL")) %>% 
  filter(!stringr::str_detect(Comments, "CELL")) %>% 
  filter(Date>"2013-12-31") %>% 
  mutate(Zipcode=as.character(Zipcode)) %>% 
  mutate(Permit_type="") %>% 
  mutate(Date=as_date(Date)) %>% 
  select(Address=Output, Zipcode, Permit_type, Permit_date=Date,
         Comments, Lon, Lat) %>% 
  filter(Permit_date<"2017-01-01")
  

generator <- rbind(generator, foo) %>% as_tibble()

charging <-  df %>% 
  filter(stringr::str_detect(Comments, "CHARGING|CHARGER")) %>% 
  filter(Permit_date>"2013-12-31") %>% 
  select(Address=Cleaned_address, Zipcode, Permit_type, Permit_date,
         Comments, Lon, Lat)

foo <- old %>% 
  filter(stringr::str_detect(Comments, "CAR CHARG")) %>% 
  filter(Date>"2013-12-31") %>% 
  mutate(Zipcode=as.character(Zipcode)) %>% 
  mutate(Permit_type="") %>% 
  mutate(Date=as_date(Date)) %>% 
  select(Address=Output, Zipcode, Permit_type, Permit_date=Date,
         Comments, Lon, Lat) %>% 
  filter(Permit_date<"2017-01-01")
  

charging <- rbind(charging, foo) %>% as_tibble()

```

## Time series

So let's look at the growth in panels over time. We'll consolidate to
monthly numbers to make the plot look descent.

The doubling time looks like about 2 years.

```{r, message=FALSE, warning=FALSE, echo=FALSE}

solar %>% 
  mutate(By_Month=floor_date(Permit_date, unit="month")) %>%
  group_by(By_Month) %>%
    summarise(MonthlyTotal=n()) %>%
  ggplot(aes(x=By_Month, y=MonthlyTotal))+
    geom_point()+
    geom_smooth() +
    labs(title="Residential Solar Panel Permits in Houston",
         subtitle="Monthly totals", 
         x="Year", 
         y="Month Total")

```

## Let's take a look at batteries

Hmmm.... looks like a big jump not long after Valentines Day of 2021.
Why would that be?

Interestingly, we see no jump in generator installs, which I really
expected to see. Which makes me a little suspicious that there is
something wrong with the data, but I don't know what that would be.

```{r, message=FALSE, warning=FALSE, echo=FALSE}

solar %>% 
  filter(str_detect(Comments, "BATT")) %>% 
  mutate(By_Month=floor_date(Permit_date, unit="month")) %>%
  group_by(By_Month) %>%
    summarise(MonthlyTotal=n()) %>%
  ggplot(aes(x=By_Month, y=MonthlyTotal))+
    geom_point()+
  annotate(geom="vline",
           x=ymd("2021-02-14"),
           xintercept=ymd("2021-02-14"),
           linetype="dashed")+
  annotate(geom="text",
           label="Feb 14, 2021",
           x=ymd("2021-02-14"),
           y=20,
           angle=90,
           vjust=-1
           ) +
    labs(title="Residential Battery Permits in Houston",
         subtitle="Monthly totals", 
         x="Year", 
         y="Month Total")


#   Generators

generator %>% 
  mutate(fuel=if_else(stringr::str_detect(Comments, " GAS "), 
                      "Gas", 
                      "Diesel")) %>% 
  mutate(as_factor(fuel)) %>% 
  mutate(By_Month=floor_date(Permit_date, unit="month")) %>%
  group_by(By_Month, fuel) %>%
    summarise(MonthlyTotal=n()) %>%
  ggplot(aes(x=By_Month, y=MonthlyTotal, color=fuel))+
    geom_point()+
    labs(title="Monthly Generator Permits in Houston",
         subtitle="Diesel installations are probably largely commercial",
         x="Year", 
         y="Month Total")

```

## Look at amount of power capacity

We will use the number of panels (where noted) to get an idea of how
much kW capacity is being added.

At the rate capacity is being added, assuming an exponential increase in
installs with a 2 year doubling time, in about 8 years the installed
capacity in the city would be the equivalent of an average Natural Gas
power plant.

```{r, message=FALSE, warning=FALSE, echo=FALSE}

panels <- solar %>% 
  mutate(Num_panels=
           as.numeric(
             str_extract(
               str_extract(Comments, "\\(\\d+\\)"),"\\d+"))) %>% 
  filter(str_detect(Comments, "RESID"),
         Num_panels<100,
         Num_panels>0)

panels %>% 
  ggplot()+
  geom_histogram(aes(x=Num_panels))

#   About 300 watts per panel

panels %>% 
  mutate(Watts=300*Num_panels/1000)  %>% 
  mutate(By_Month=floor_date(Permit_date, unit="month")) %>%
  group_by(By_Month) %>%
    summarise(MonthlyTotal=sum(Watts)) %>%
  ggplot(aes(x=By_Month, y=MonthlyTotal))+
    geom_point()+
    geom_smooth() +
    labs(title="Monthly Residential Solar Panel Kilowatts in Houston",
         subtitle="For installations where the number of panels was noted", 
         x="Year", 
         y="Month Total")

#   Assume average 30 panels per install, at 300 watts per, so 9 kwatts per

solar %>% 
  mutate(By_Month=floor_date(Permit_date, unit="month")) %>%
  group_by(By_Month) %>%
    summarise(MonthlyTotal=n()) %>%
  mutate(Kwatts=MonthlyTotal*9) %>% 
  ggplot(aes(x=By_Month, y=Kwatts))+
    geom_point()+
    geom_smooth() +
    labs(title="Residential Solar Panel Total kW added in Houston",
         subtitle="Monthly totals", 
         x="Year", 
         y="Month Total")

```

## Let's look at the distribution by Zip code

Surprisingly the Houston Arrow does not appear on this map. That is a
bit of a surprise. Note that 77048 is dominated by a property of
apartment homes that were built with a large solar array.

```{r, message=FALSE, warning=FALSE, echo=FALSE}

#   Top 20 zipcodes

solar %>% 
  group_by(Zipcode) %>% 
  summarise(n = n()) %>% 
  dplyr::arrange(desc(n)) %>%
  dplyr::slice(1:20) %>% 
  gt::gt() %>% 
  gt::tab_header(
    title = "Solar Panel Installations by Zip Code",
    subtitle = "2014 - present"
  ) %>% 
  gt::cols_label(
    Zipcode = "Zip Code",
    n = "Number of Installations"
  )

#######    now let's make a map

zippath="/home/ajackson/Dropbox/Rprojects/Curated_Data_Files/Zipcodes/"
zips <- readRDS(paste0(zippath, "COH_Zip_Polys.rds")) %>% 
  mutate(Zipcode=as.character(ZIP_CODE)) %>% 
  sf::st_as_sf()


#   Number of permits per zip

solar_zip <- solar %>% 
  group_by(Zipcode) %>% 
  summarise(Num = n()) %>% 
  mutate(Label=paste(as.character(Num) , "permits", Zipcode)) %>%
  mutate(Num=cut_width(Num, 25))

solar_zip <- right_join(zips, solar_zip, by="Zipcode")

# Create a factored palette function
pal <- colorFactor(
  palette = "Blues",
  domain = solar_zip$Num)

leaflet::leaflet(solar_zip) %>% 
  setView(lng = -95.362306, lat = 29.756931, zoom = 12) %>% 
  addTiles() %>%
  addPolygons(weight=1,
              fillColor = ~pal((Num)),
              fillOpacity = 0.5,
              label = ~Label) %>% 
  addLegend("bottomleft", pal = pal, values = ~Num,
    title = "Num Permits",
    opacity = 1
  )

```

## Let's combine solar panels with assessments

Interestingly, it is clear that solar panels are not just for the
wealthy. The median value of homes with panels is pretty close to the
median value of homes in general. I used the 2021 HCAD total market
value for the value for each home.

There are some differences however. It appears that it roughly twice as
likely that a house worth over \$1,000,000 will have solar panels than a
house worth less than that.

```{r, message=FALSE, warning=FALSE, echo=FALSE}

HCADpath="/home/ajackson/Dropbox/Rprojects/Curated_Data_Files/HCAD/"
HCAD <- readRDS(paste0(HCADpath, "Values_2021.rds")) %>% 
  rename(Address=site_addr_1)

foo <- inner_join(solar, HCAD, by="Address") %>% 
  mutate(Value=as.numeric(tot_mkt_val)) %>% 
  filter(!stringr::str_detect(state_class, "X")) %>% 
  select(Address, Zipcode, Permit_date, Comments, new_own_dt, Value, state_class)

foo %>% 
  filter(Value<2000000) %>% 
  ggplot(aes(x=Value)) +
  geom_histogram(bins=32)+
  scale_x_continuous(labels=scales::dollar_format()) +
  labs(title="Property Market Value",
       subtitle="For Properties with Solar Panels",
       x="Market Value",
       y="Number of Properties")

HCAD %>% 
  mutate(Value=as.numeric(tot_mkt_val)) %>% 
  filter(!stringr::str_detect(state_class, "X")) %>% 
  filter(Value<2000000) %>% 
  ggplot(aes(x=Value)) +
  geom_histogram(bins=32)+
  scale_x_continuous(labels=scales::dollar_format()) +
  labs(title="Property Market Value",
       subtitle="For All Properties",
       x="Market Value",
       y="Number of Properties")


foo2 <- HCAD %>% 
  mutate(Value=as.numeric(tot_mkt_val)) %>% 
  filter(!stringr::str_detect(state_class, "X")) %>% 
  filter(Value<2000000) %>% 
  mutate(Bins=cut_width(Value, 100000)) %>% 
  group_by(Bins) %>% 
    summarize(n=n())

foo %>% 
  filter(Value<2000000) %>% 
  mutate(Bins=cut_width(Value, 100000)) %>% 
  group_by(Bins) %>% 
    summarize(nsolar=n()) %>% 
  left_join(., foo2, by="Bins") %>% 
  mutate(Pct=nsolar/n) %>%
  cbind(., seq(0000, 2050000, 100000)) %>% 
  rename(Value=5) %>% 
  ggplot(aes(x=Value, y=Pct)) +
  geom_histogram(bins=20, stat="identity")+
  scale_x_continuous(labels=scales::dollar_format()) +
  scale_y_continuous(labels=scales::percent) +
  labs(title="Property Market Value",
       subtitle="For All Properties",
       x="Market Value",
       y="Percent of Properties")
  
  
  
```

## Car charging

One final somewhat related topic - electric car chargers. These are all
commercial, I don't think residential chargers (220 volt) require a
permit.

```{r, message=FALSE, warning=FALSE, echo=FALSE}

charging %>% 
  mutate(By_Month=floor_date(Permit_date, unit="month")) %>%
  group_by(By_Month) %>%
    summarise(MonthlyTotal=n()) %>%
  ggplot(aes(x=By_Month, y=MonthlyTotal))+
    geom_point()+
    labs(title="Electric Car Charger Permits in Houston",
         subtitle="Monthly totals", 
         x="Year", 
         y="Month Total")

```
