---
title: "Parking Lot Study"
author: "Alan Jackson"
date: "2022-12-06"
slug: parking-lot-study
categories: 
- Infrastructure
tags:
- Houston
- Geostatistics
- Mapping
keywords: tech
comments: no
showMeta: no
showActions: no
output:
  html_document:
    code_folding: hide
---

##    Parking Lots, Parking Minimums, and All That

In 2019, Houston city council passed a ground-breaking ordinance that removed
parking minimums for parts of Midtown and EaDo. The city’s climate action plan,
which was adopted in 2020, calls for an end to all minimum parking requirements
by 2030.

I thought it would be interesting to look at parking in The Heights, so I put
together this little study. Originally I wanted to do the entire city, but I
could not find a source of data on parking lots covering the whole city, so
that was out. For this study I used OpenStreetMap, and added polygons for the
parking lots in the west Heights. The editor is pretty easy to use, but that was
still quite a few hours of work. I basically cover the area delineated by 20th
street to the north, 11th street to the south, Durham to the west, and Heights
Boulevard to the east.

One item of interest in the Heights is that the area along 19th street, which is
largely businesses, has no parking minimums because it was built prior to those
requirements. We'll look at the different business corridors to see how they
compare.

I'm also interested in how many people have easy access to businesses, either by
walking or bicycling, and also how much potential property tax revenue is being
lost by covering land with parking lots instead of buildings.

A good overview of many of the issues may be found in [this
paper](https://kinder.rice.edu/urbanedge/less-space-parking-crucial-step-toward-walkability-houston)
from the Kinder Institute.

```{r setup, include=FALSE, echo=FALSE}

library(tidyverse)
library(GeocodeHou)
library(postmastr)

google_crs <- "EPSG:4326" # lat long
CoH_crs <- "EPSG:2278" # X-Y

knitr::opts_chunk$set(echo = TRUE)
```

##   Read in the necessary datasets


###  First let's establish an AOI 

The buffer is defined along the streets of interest, and the AOI is 50% larger
to ensure we grab the needed data.

```{r AOI, message=FALSE, warning=FALSE}
###{r initialize, class.source = 'fold-show'}

#######################################
#   AOI
#######################################

#   Start at 19th and Shepherd, east to Yale, south to 11th, west to Shepherd
Buffer_coord <- as_tibble(rbind(c(29.802783, -95.410146),
                      c(29.802935, -95.399204),
                      c(29.790674, -95.398971),
                      c(29.790514, -95.409865),
                      c(29.802783, -95.410146)))

Study_box <- Buffer_coord %>%
  sf::st_as_sf(coords = c("V2", "V1"), crs = google_crs) %>%
  summarise(geometry = sf::st_combine(geometry)) %>%
  sf::st_cast("LINESTRING")

AOI <- Study_box %>% #  enlarge by 15,000 feet
  sf::st_transform(crs=CoH_crs) %>% 
  sf::st_line_sample(n=250) %>% 
  sf::st_buffer(dist=10000, endCapStyle="SQUARE")

bbox <- sf::st_bbox(sf::st_transform(lines, crs=CoH_crs))
#AOI <- bbox

#bbox <- expand_bb(bbox, 0.5)
#AOI <- expand_bb(AOI, 2)


#   Create a ggplot basemap object

Base_basemapR <- basemapR::base_map(sf::st_bbox(
                                      sf::st_transform(AOI, crs=google_crs)),
                                    basemap="mapnik", increase_zoom=2)
ggplot() +
  Base_basemapR +
  geom_sf(data=sf::st_transform(AOI, crs=google_crs), aes(color="red"), alpha=0.0)+
  geom_sf(data=Study_box, aes(fill="red"), alpha=0.3)#+
  # ggtitle("Total Daytime EMS Calls per Keymap Square",
  #         subtitle=Time_period)

leaflet::leaflet() %>%
  leaflet::addTiles() %>% # OpenStreetMap by default
  leaflet::addPolylines(data=sf::st_transform(Study_box, crs=google_crs),
                   opacity=0.5,
                   fillOpacity = 0.3) %>% 
  leaflet::addPolygons(data=sf::st_transform(sf::st_as_sfc(bbox), crs=google_crs),
                   color="red",
                   weight=2,
                   opacity=0.5,
                   fillOpacity = 0.3)   %>% 
  leaflet::addPolygons(data=sf::st_transform(AOI, crs=google_crs),
                   color="green",
                   weight=2,
                   opacity=0.5,
                   fillOpacity = 0.3)   
```

###     Load parking lots

From OpenStreetMap we will load the digitized parking lots that I digitized
for the study.

```{r load parking lots, message=FALSE, warning=FALSE}
#######################################
#   Parking lots
#######################################

parking <- 
  osmdata::getbb(place_name = "Houston") %>% 
  osmdata::opq(timeout = 50) %>% # Build query
  osmdata::add_osm_feature(key = "amenity", 
                           value = "parking" 
                            ) %>% 
  osmdata::osmdata_sf(quiet=FALSE) # turn into an sf file

#   Extract parking lots in the AOI

Parking_TF <- sf::st_contains( sf::st_transform(
                                    sf::st_as_sfc(bbox, crs=google_crs),
                                    crs=CoH_crs),
                               sf::st_transform(
                                    sf::st_as_sfc(parking$osm_polygons,
                                              crs=google_crs),
                                    crs=CoH_crs))

Parking_bb <- parking$osm_polygons[unlist(Parking_TF),]

leaflet::leaflet() %>%
  leaflet::addTiles() %>% # OpenStreetMap by default
  leaflet::addPolygons(data=Parking_bb,
                   opacity=0,
                   fillOpacity = 1)
```



```{r  load census data, message=FALSE, warning=FALSE}
#######################################
#   Census data
#######################################

path <- "/home/ajackson/Dropbox/Rprojects/Curated_Data_Files/Census_data/"


#######################################
#   HCAD data
#######################################


```









