---
title: "Parking Lot Study"
author: "Alan Jackson"
date: "2022-12-06"
slug: parking-lot-study
categories: 
- Infrastructure
tags:
- Houston
- Geostatistics
- Mapping
keywords: tech
comments: no
showMeta: no
showActions: no
output:
  html_document:
    code_folding: hide
---

##    Parking Lots, Parking Minimums, and All That

In 2019, Houston city council passed a ground-breaking ordinance that removed
parking minimums for parts of Midtown and EaDo. The city’s climate action plan,
which was adopted in 2020, calls for an end to all minimum parking requirements
by 2030.

I thought it would be interesting to look at parking in The Heights, so I put
together this little study. Originally I wanted to do the entire city, but I
could not find a source of data on parking lots covering the whole city, so
that was out. For this study I used OpenStreetMap, and added polygons for the
parking lots in the west Heights. The editor is pretty easy to use, but that was
still quite a few hours of work. I basically cover the area delineated by 20th
street to the north, 11th street to the south, Durham to the west, and Heights
Boulevard to the east.

One item of interest in the Heights is that the area along 19th street, which is
largely businesses, has no parking minimums because it was built prior to those
requirements. We'll look at the different business corridors to see how they
compare.

I'm also interested in how many people have easy access to businesses, either by
walking or bicycling, and also how much potential property tax revenue is being
lost by covering land with parking lots instead of buildings.

A good overview of many of the issues may be found in [this
paper](https://kinder.rice.edu/urbanedge/less-space-parking-crucial-step-toward-walkability-houston)
from the Kinder Institute.

```{r setup, include=FALSE, echo=TRUE}

library(tidyverse)
library(GeocodeHou) # personal package for geocoding
library(postmastr) # nice address parsing package
library(tmap)
library(tmaptools)
library(sf)

google_crs <- "EPSG:4326" # lat long
CoH_crs <- "EPSG:2278" # X-Y

knitr::opts_chunk$set(echo = TRUE)
```

##   Read in the necessary datasets


###  First let's establish an AOI_xy 

The buffer is defined along the streets of interest, and the AOI_xy is 10,000
feet larger to ensure we grab the needed data.

```{r AOI_xy, message=FALSE, warning=FALSE}
###{r initialize, class.source = 'fold-show'}

#######################################
#   AOI_xy
#######################################

#   Start at 19th and Shepherd, east to Yale, south to 11th, west to Shepherd

Buffer_coord <- tribble(~Lat,      ~Lon,
                        29.802783, -95.410146,
                        29.802935, -95.399204,
                        29.790674, -95.398971,
                        29.790514, -95.409865,
                        29.802783, -95.410146)

Study_box_ll <- Buffer_coord %>%
  sf::st_as_sf(coords = c("Lon", "Lat"), crs = google_crs) %>%
  summarise(geometry = sf::st_combine(geometry)) %>%
  sf::st_cast("LINESTRING")

AOI_xy <- Study_box_ll %>% #  enlarge by 10,000 feet
  sf::st_transform(crs=CoH_crs) %>% 
  sf::st_line_sample(n=250) %>% 
  sf::st_buffer(dist=10000, endCapStyle="SQUARE")

AOI_ll <- AOI_xy %>% sf::st_transform(crs=google_crs)

#   Buffer used to extract parking lots for the study

Parking_select_xy <- Study_box_ll %>% #  enlarge by 1000 feet
  sf::st_transform(crs=CoH_crs) %>% 
  sf::st_line_sample(n=250) %>% 
  sf::st_buffer(dist=1000, endCapStyle="ROUND")

bbox_xy <- sf::st_bbox(sf::st_transform(Study_box_ll, crs=CoH_crs))
bbox_ll <- sf::st_bbox(Study_box_ll)
bbox_AOI_ll <- sf::st_bbox(AOI_ll)

#   Set default basemap
tmap::tmap_options(basemaps="OpenStreetMap")

tmap::tmap_mode("view") # set mode to interactive plots

tmap::tm_shape(sf::st_as_sf(as_tibble(AOI_ll) %>% mutate(Label="AOI"))) +
  tmap::tm_polygons(alpha=0.3, col="blue")+
tmap::tm_shape(sf::st_as_sf(as_tibble(sf::st_cast(Study_box_ll, "POLYGON")) %>% mutate(Label="Study area"))) +
  tmap::tm_polygons(alpha=0.3, col="red")+
  tmap::tm_scale_bar()+
  tmap::tmap_options(unit = "mi") +
  tmap::tm_layout(title="Area of Interest and Study Area")

```

##        I need a freeway polygon

It is painful for pedestrians and bicyclists to cross under the freeway. The
roads pretty effectively segregate the neighborhood. I will honor that by
clipping my distance polygons with the freeways that basically define the
boundaries of the Heights.

```{r make freeway poly, cache=FALSE}

#   Get freeway data

if (!file.exists("OSM_Freeways.rds")) { # save results
  Freeways <- 
    osmdata::getbb(place_name = "Houston") %>% 
    osmdata::opq(timeout=50)%>% # Build query
    osmdata::add_osm_feature(key = "highway", 
                    value = c("motorway", 
                              "motorway_link" 
                              )) %>% # select the big roads
    osmdata::osmdata_sf(quiet=FALSE) 
  
  saveRDS(Freeways, "OSM_Freeways.rds")
} else {
  Freeways <- readRDS("OSM_Freeways.rds")
}


#   These are the freeways that will define a polygon
Fwy_names <- paste0(c("North Loop", "Katy Freeway", "West Loop North", 
               "North Freeway", "North Loop West"), collapse="|")

Selected_fwys <- Freeways$osm_lines %>% 
  filter(stringr::str_detect(name, Fwy_names)) %>% 
  select(osm_id, name)

#   Create polygons and (interactively) figure out which one is wanted

Fwy_poly <- Selected_fwys %>% 
  st_union() %>% 
  st_polygonize() %>% 
  st_cast() %>% 
  as_tibble() %>% 
  mutate(id=row_number()) %>% 
  st_as_sf() %>% 
  filter(id==5)

#     Let's take a look

tmap::tmap_options(basemaps="OpenStreetMap")
tmap::tmap_mode("view") # set mode to interactive plots

tmap::tm_shape(sf::st_as_sf(as_tibble(Fwy_poly) %>% mutate(Label="Freeways"))) +
  tmap::tm_polygons(alpha=0.3, col="blue")+
  tmap::tm_scale_bar()+
  tmap::tmap_options(unit = "mi")# +
  # tmap::tm_layout(title="Freeway-defined Study Area Cutoff")

```

##        Collect the data

First we will collect the various data we will need in the analysis. 

- Parking lots
- Census data
- Tax data

###     Load parking lots

From OpenStreetMap we will load the digitized parking lots that I digitized
for the study.

```{r load parking lots, message=FALSE, warning=FALSE, cache=FALSE}
#######################################
#   Parking lots
#######################################

if (!file.exists("OSM_parking.rds")) { # save results
  parking <- 
    osmdata::getbb(place_name = "Houston") %>% 
    osmdata::opq(timeout = 50) %>% # Build query
    osmdata::add_osm_feature(key = "amenity", 
                             value = "parking" 
                              ) %>% 
    osmdata::osmdata_sf(quiet=FALSE) # turn into an sf file
  
  saveRDS(parking, "OSM_parking.rds")
} else {
  parking <- readRDS("OSM_parking.rds")
}

#   Extract parking lots in the AOI that are near the study box 

Parking_TF <- sf::st_contains( Parking_select_xy, 
                               sf::st_transform(
                                    sf::st_as_sfc(parking$osm_polygons,
                                              crs=google_crs),
                                    crs=CoH_crs))

Parking_ll <- parking$osm_polygons[unlist(Parking_TF),] %>% 
  select(osm_id, amenity, surface)


tmap::tmap_options(basemaps="OpenStreetMap")
tmap::tmap_mode("view") # set mode to interactive plots

tmap::tm_shape(Parking_ll) +
  tmap::tm_polygons(alpha=0.8, col="blue")+
  tmap::tm_scale_bar()+
  tmap::tmap_options(unit = "mi")# +
  # tmap::tm_layout(title="Parking lots")

```

###       Load census data

I have previously downloaded and tidied census data for this area, so I will
utilize that.

I'll load the population by block data, and the income by block-group data.
I'll also calculate some useful stuff like population density.


```{r  load census data, message=FALSE, warning=FALSE}
#######################################
#   Census data
#######################################

path <- "/home/ajackson/Dropbox/Rprojects/Curated_Data_Files/Census_data/"

Pop <- readRDS(paste0(path, "Pop_data_by_block_2020.rds"))
ACS <- readRDS(paste0(path, "Income_and_Age_by_BlkGrp_2015_2020.rds"))

#   Extract census blocks in the AOI that are near the study box 

Pop_TF <- sf::st_contains(AOI_xy, 
                          sf::st_transform(Pop, crs=CoH_crs)) 

Pop_ll <- Pop[unlist(Pop_TF),]  %>% 
  sf::st_cast("POLYGON") %>% 
  sf::st_as_sf()

tmap::tmap_options(basemaps="OpenStreetMap")
tmap::tmap_mode("view") # set mode to interactive plots

tmap::tm_shape(Pop_ll) +
  tmap::tm_polygons(alpha=0.8, col="Pop", 
                    breaks = c(100, 200, 300, 400, 500, 1000))+
  tmap::tm_scale_bar()+
  tmap::tmap_options(unit = "mi")# +
  # tmap::tm_layout(title="Population")

#   Calculate Population density

Pop_ll <- 
  Pop_ll %>%
  mutate(area=sf::st_area(geometry)) %>% 
  mutate(area=units::drop_units(area)) %>% 
  mutate(Pop_density=(Pop/area)*2.59e+6) %>% 
  mutate(Pop_density=ifelse(Pop<10, NA, Pop_density))

tmap::tm_shape(Pop_ll %>% 
                 mutate(Pop_density=if_else(is.na(Pop_density),0,Pop_density))) +
  tmap::tm_polygons(alpha=0.8, col="Pop_density", 
                    breaks = c(0, 5000, 10000, 20000, 40000))+
  tmap::tm_scale_bar()+
  tmap::tmap_options(unit = "mi")# +
  # tmap::tm_layout(title="Population Density (per/sq-mi)")

#   Income

#   Extract census block groups in the AOI that are near the study box 

ACS_TF <- sf::st_contains(AOI_xy, 
                          sf::st_transform(ACS, crs=CoH_crs)) 

ACS_ll <- ACS[unlist(ACS_TF),]  %>% 
  sf::st_cast("POLYGON") %>% 
  sf::st_as_sf()

tmap::tm_shape(ACS_ll) +    
  tmap::tm_polygons(alpha=0.8, col="Per_cap_incE", 
                    breaks = c(0, 50000, 100000, 200000, 400000))+
  tmap::tm_scale_bar()+
  tmap::tmap_options(unit = "mi") #+
  # tmap::tm_layout(title="Average Income")

```


```{r load Tax data}

#######################################
#   HCAD data
#######################################

Hcad_path <- "/home/ajackson/Dropbox/Rprojects/Curated_Data_Files/"
df <- readRDS(paste0(Hcad_path, "HCAD/Values_2022.rds"))

#   First pare down file by zipcode

df <- df %>% 
  filter(stringr::str_detect(site_addr_3,
                             paste(c("77008", "77007", "77009"), collapse="|")))

#   Pull out just the commercial properties

df <- df %>% 
  filter(state_class=="F1")

#   Calculate some stuff

df <- df %>% 
  mutate(bld_ar=as.numeric(bld_ar)) %>% 
  mutate(land_ar=as.numeric(land_ar)) %>% 
  mutate(bld_val=as.numeric(bld_val)) %>% 
  mutate(land_val=as.numeric(land_val)) %>% 
  mutate(ratio_area=bld_val/land_val) %>% 
  mutate(ratio_value=bld_val/land_val) %>% 
  mutate(bld_persqft=bld_val/bld_ar) %>% 
  mutate(land_persqft=land_val/land_ar) %>% 
  filter(bld_ar>500) %>% 
  filter(land_ar>1000) %>% 
  filter(land_val>0)
  

#   How big are the commercial buildings?

df %>% 
  ggplot(aes(x=bld_ar, fill=site_addr_3)) +
  geom_histogram() +
  labs(title="Building Areas in Zips 77007, 77008, and 77009",
       x="Area (sq ft)")

#   Value per sq foor for buildings

df %>% 
  ggplot(aes(x=bld_persqft, fill=site_addr_3)) +
  geom_histogram() +
  labs(title="Building Value per Square Foot in Zips 77007, 77008, and 77009",
       x="Dollars per sq-ft")

#   Value per sq foor for land

df %>% 
  ggplot(aes(x=land_persqft, fill=site_addr_3)) +
  geom_histogram(bins=20) +
  labs(title="Land Value per Square Foot in Zips 77007, 77008, and 77009",
       x="Dollars per sq-ft")

```









