---
title: Texas Deaths by County
author: Alan Jackson
date: '2022-05-22'
slug: '[texas_deaths_by_county]'
categories:
  - Mortality
tags:
  - Texas
keywords:
  - tech
---

```{r setup, include=FALSE}

library(tidyverse)
library(ggplot2)
library(lubridate)

#   Directory where data is stored

DataCovid <- "/home/ajackson/Dropbox/Rprojects/Covid/Today_Data/"
DataVotes <- "/home/ajackson/Dropbox/Rprojects/Voting/"
DataDeaths <- "/home/ajackson/Dropbox/Rprojects/Curated_Data_Files/CDC_Wonder/"

#   Build Votes by county tibble

Votes <- read_csv(paste0(DataVotes, 
                         "CountybyCountyCanvassReport2020Pres.csv")) %>% 
  rename(Candidate="CANDIDATE NAME", 
         "Office"="OFFICE NAME", 
         "County"="COUNTY NAME",
         Votes="TOTAL VOTES PER OFFICE PER COUNTY") %>% 
  filter(grepl("BIDEN|TRUMP", Candidate)) %>% 
  select(Candidate, County, Votes) %>% 
  pivot_wider(names_from=Candidate, values_from=Votes) %>% 
  rename(Biden=2, Trump=3) %>% 
  mutate(Blueness=Biden/(Biden+Trump))

Covid <- read_rds(paste0(DataCovid, "Today_County_calc.rds"))

out_path <-  "/home/ajackson/Dropbox/Rprojects/Covid/"
Vaccine <- read_rds(paste0(out_path, "Vaccinations.rds")) %>% 
  filter(County!="NULL", 
         County!="Other", 
         County!="*Other", 
         County!="* Other", 
         County!="Texas", 
         !is.na(County),
         County != "Federal Long-Term Care Vaccination Program",
         County != "Federal Pharmacy Retail Vaccination Program") %>% 
  group_by(County) %>% 
    mutate(Pop_teen=max(Pop_teen, na.rm=TRUE)) %>% 
  ungroup()

#   Add total population to file

Vaccine <- Covid %>% 
  select(County, Population) %>% 
  unique() %>% 
  rename(Pop_total=Population) %>% 
  right_join(., Vaccine, by="County")

today_date <- stamp("Aug 1, 2021")(last(Vaccine$Date))

Deaths <- readRDS(paste0(DataDeaths,
                         "Texas_Deaths_Jan2018-Mar2022.rds")) %>% 
  mutate(Year=factor(lubridate::year(Month)),
         Year=lubridate::year(Month),
         Date=Month,
         Month=lubridate::month(Month, label=TRUE, abbr=TRUE)
         ) %>% 
  rename(County= `Residence County`) %>% 
  mutate(County=stringr::str_remove(County, " County, TX"),
         County=stringr::str_to_upper(County))

```

##   Covid and Politics

Let's take a look at Texas politics and Covid deaths. The CDC Wonder data
has preliminary deaths and preliminary covid deaths, by county through April
of 2022. We can combine this data with the votes from the presidential race
and look for correlations.

But first, I will calculate the "excess" deaths using a bog simple approach,
I will simply assume that for each county the deaths in years 2018 and 2019
represent a somewhat constant background, and any increase in the years 2020 
and 2021 will be considered excess deaths likely due to Covid.

In the plots, I look at the nine counties with the largest per capita excess
between 2018 and 2021. The data is pretty noisy, especially since these are
small counties, so the statistics are not very robust.Even so, the picture
that emerges is pretty consistent.

```{r plots, message=FALSE, warning=FALSE}

#   Look at year to year differences compared to 2018

df <- Deaths %>% 
  add_count(County) %>% 
  filter(n==51) %>% 
  filter(Year<2022) %>% 
  pivot_wider(c(County, Month), names_from=Year, values_from = Deaths) %>% 
  rename(y2018=`2018`, y2019=`2019`, y2020=`2020`, y2021=`2021`) %>% 
  mutate(y2018=replace_na(y2018, 0),
         y2019=replace_na(y2019, 0),
         y2020=replace_na(y2020, 0),
         y2021=replace_na(y2021, 0)) %>% 
  mutate(Delta_1=y2019-y2018,
         Delta_2=y2020-y2018,
         Delta_3=y2021-y2018
         ) %>% 
  group_by(County) %>% 
    summarize(n = n(),
              D_1=mean(Delta_1, na.rm=TRUE)*12,
              D_2=mean(Delta_2, na.rm=TRUE)*12,
              D_3=mean(Delta_3, na.rm=TRUE)*12) %>% 
  mutate(Total=D_2+D_3) 
  
Pops <- Covid %>% 
  select(County, Population) %>% 
  unique() %>% 
  mutate(County=stringr::str_to_upper(County))

df <- left_join(df, Pops, by="County")

df <- df %>% 
  mutate(D_1=D_1/Population,
         D_2=D_2/Population,
         D_3=D_3/Population,
         Totalpercap=Total/Population)

Top_9 <- df %>% 
  arrange(-Totalpercap) %>% 
  head(9)

Deaths %>% 
  filter(Year<2022) %>% 
  filter(County %in% Top_9$County) %>% 
  mutate(Year=factor(Year)) %>% 
  ggplot(aes(x=Month, y=Deaths, group=Year, color=Year)) +
  geom_point() +
  geom_line(aes(x=Month)) +
  facet_wrap(vars(County), scales="free")+
  theme_classic()

```


##  Excess deaths


Let's do a very simple analysis. We will look at the number of deaths, per
county, in 2018 and 2019, and compare that to 2020 and 2021. We will call that
excess deaths. 

The histograms clearly shows that almost every county experienced an
excess of deaths, most on the order of 500 per 100,000 in those two years.
Clearly there is a strong signal here.

```{r excess deaths, message=FALSE, warning=FALSE, echo=FALSE}

Yearly_deaths <- read_tsv(paste0(DataDeaths,
                "Provisional Mortality Statistics, 2018-2021 by year.txt"),
                col_types="-cccci") %>% 
  mutate(Year=factor(`Year Code`)
         ) %>% 
  rename(County= `Residence County`) %>% 
  mutate(County=stringr::str_remove(County, " County, TX"),
         County=stringr::str_to_upper(County))

Excess_deaths <-  Yearly_deaths %>% 
  select(County, Year, Deaths) %>% 
  drop_na() %>% 
  add_count(County) %>% 
  filter(n==4) %>% 
  pivot_wider(c(County), names_from=Year, values_from=Deaths) %>% 
  mutate(Pre_covid=(`2018`+`2019`),
         Post_covid=(`2020`+`2021`),
         Excess=Post_covid-Pre_covid) %>% 
  left_join(Pops, by="County") %>% 
  mutate(Excess_percap=Excess/Population*100000)

Excess_deaths %>% 
  ggplot(aes(x=Excess_percap)) +
  geom_histogram() +
  labs(title="Excess Deaths per Capita by County",
       subtitle="2018 and 2019 compared to 2020 and 2021 (2 year total)",
       x="Excess deaths per 100,000",
       y="Number of counties")

```

##  Add in CDC Covid estimates

So the CDC has preliminary numbers for Covid deaths based on death certificate
data, although we know that in some counties (where the coroner is a JP) those
numbers are low. One JP was quoted as saying "why would I waste a Covid test 
on a dead person?".

None the less, their data shows a rate of close to 400 per 100,000, and less
spread than the excess death estimate. 

```{r covid, message=FALSE, warning=FALSE, echo=FALSE}
Covid_deaths <- read_tsv(paste0(DataDeaths,
                "Provisional Mortality Statistics, 2018-2021 covid.txt"),
                col_types="-cccci") %>% 
  mutate(Year=factor(`Year Code`)
         ) %>% 
  rename(County= `Residence County`) %>% 
  mutate(County=stringr::str_remove(County, " County, TX"),
         County=stringr::str_to_upper(County))

Covid_deaths <- 
Covid_deaths %>% 
  drop_na() %>% 
  select(County, Year, Deaths) %>% 
  group_by(County) %>% 
    summarize(Covid=sum(Deaths, na.rm=TRUE)) %>% 
  left_join(Pops, by="County") %>% 
  mutate(Covid_percap=Covid/Population*100000)

Covid_deaths %>% 
  ggplot(aes(x=Covid_percap)) +
  geom_histogram() +
  labs(title="Covid Deaths per Capita by County",
       subtitle="Preliminary CDC estimates",
       x="Excess deaths per 100,000",
       y="Number of counties")


```


##  Compare CDC covid deaths and excess deaths

When we compare the two estimates, we find that there appears to be about
a 150 deaths per 100,000 difference, with the excess deaths being fairly
consistently larger than the proven Covid deaths. In light of what we know
about the construction of death certificates, this is not surprising.

```{r compare, message=FALSE, warning=FALSE, echo=FALSE}

foo <- left_join(Covid_deaths, Excess_deaths, by="County") %>% 
  select(County, Covid, Covid_percap, 
         Excess, Excess_percap, Population=Population.y) %>% 
  mutate(Excess_minus_Covid=(Excess-Covid)/Population*100000)

foo %>% 
  ggplot(aes(x=Excess_minus_Covid))+
  geom_histogram()+
  labs(title="(Excess - Covid) Deaths per Capita by County",
       x="Deaths per 100,000",
       y="Number of counties")

```


##        Analysis of Excess vs Covid deaths

Excess and Covid deaths are clearly correlated, but fairly noisy. But more
importantly, the slope of the linear fit of that correlation is only 0.3,
which implies that not only is there a constant shift from Covid deaths to
Excess deaths, but that the shift itself has a bias - the counties with
the largest excess deaths are possibly under-reporting their Covid deaths.
Or counties with the lowest excess deaths are over-reporting Covid deaths, 
which seems unlikely. And this is not a small bias.

```{r excess vs covid analysis, message=FALSE, warning=FALSE}

#   Covid vs. Excess

foo %>% 
  ggplot(aes(x=Excess_percap, y=Covid_percap)) +
  geom_point()+
  geom_smooth(method=lm)+
  annotate("text", x=250, y=700, 
           label=(paste0("slope==",
                         signif(coef(lm(foo$Covid_percap~foo$Excess_percap))[2],3))),
           parse=TRUE)+
  labs(title="Excess Deaths vs. Covid Deaths (per 100,000)",
       subtitle="By County",
       x="Excess Deaths per 100,000",
       y="Covid Deaths per 100,000")

#   County pop vs. Covid, Excess, and difference

p1 <- foo %>% 
  ggplot(aes(x=Population, y=Covid_percap))+
  geom_point()+
  scale_x_log10() +
  geom_smooth(method=lm)+
  labs(title="Covid Deaths (per 100,000) vs. Population",
       subtitle="By County",
       x="Population",
       y="Covid Deaths per 100,000")
  

p2 <- foo %>% 
  ggplot(aes(x=Population,y=Excess_percap ))+
  geom_point()+
  scale_x_log10()+
  geom_smooth(method=lm)+
  labs(title="Excess Deaths (per 100,000) vs. Population",
       subtitle="By County",
       x="Population",
       y="Excess Deaths per 100,000")

p3 <- foo %>% 
  ggplot(aes(x=Population, y=Excess_minus_Covid))+
  geom_point()+
  scale_x_log10()+
  geom_smooth(method=lm)+
  labs(title="Excess - Covid Deaths (per 100,000) vs. Population",
       subtitle="By County",
       x="Population",
       y="Death differences")

gridExtra::grid.arrange(p1, p2, p3, nrow=3)

```

##  Add in politics

```{r politics}

By_county <- left_join(foo, Votes, by="County") %>% 
  select(-Biden, -Trump)

#   First Trauma Service Areas 

inpath <- "/home/ajackson/Dropbox/Rprojects/Curated_Data_Files/Texas_Trauma_Service_Areas/"

TSA <- readRDS(paste0(inpath, "Trauma_Service_Areas.rds"))

By_TSA <- TSA %>% 
  mutate(County=stringr::str_to_upper(County)) %>% 
           left_join(foo, ., by="County") %>% 
           left_join(., Votes, by="County") %>% 
  group_by(Area_name) %>% 
    summarize(Population=sum(Population, na.rm = TRUE),
              Covid=sum(Covid, na.rm=TRUE),
              Excess=sum(Excess, na.rm=TRUE),
              Trump=sum(Trump, na.rm=TRUE),
              Biden=sum(Biden, na.rm=TRUE)
              ) %>% 
  mutate(Blueness=Biden/(Trump+Biden)) %>% 
  rename(TSA=Area_name)


#   Metropolitan areas

DataArchive <- "/home/ajackson/Dropbox/mirrors/ajackson/SharedData/"
MSA_raw <- readRDS(paste0(DataArchive, "Texas_MSA_Pop_Counties.rds")) 

By_MSA <- MSA_raw %>% 
  unnest(Counties) %>% 
  rename(County=Counties) %>% 
  mutate(County=stringr::str_to_upper(County)) %>% 
  left_join(foo, ., by="County") %>% 
  left_join(., Votes, by="County") %>% 
  select(-Population.x) %>% 
  rename(Population=Population.y) %>% 
  group_by(MSA) %>% 
    summarise(Population=unique(Population),
              Covid=sum(Covid, na.rm=TRUE),
              Excess=sum(Excess, na.rm=TRUE),
              Trump=sum(Trump, na.rm=TRUE),
              Biden=sum(Biden, na.rm=TRUE)
            ) %>% 
  mutate(Blueness=Biden/(Trump+Biden)) 
  

```

##   Let's make some plots

```{r political plots}

#  By County

By_county %>% 
  ggplot(aes(x=Blueness,y=Excess_percap ))+
  geom_point()+
  geom_smooth(method=lm)+
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(title="Excess Deaths (per 100,000) vs. Biden percent vote",
       subtitle="By County",
       x="Biden Percent Vote",
       y="Excess Deaths per 100,000")

#  By MSA

By_MSA %>% 
  mutate(Excess_percap=Excess/Population*100000) %>% 
  ggplot(aes(x=Blueness,y=Excess_percap ))+
  geom_point()+
  geom_smooth(method=lm)+
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  labs(title="Excess Deaths (per 100,000) vs. Biden percent vote",
       subtitle="By Metropolitan Area",
       x="Biden Percent Vote",
       y="Excess Deaths per 100,000")

#  By TSA

By_TSA %>% 
  mutate(Excess_percap=Excess/Population*100000) %>% 
  ggplot(aes(x=Blueness,y=Excess_percap ))+
  geom_point()+
  geom_smooth(method=lm)+
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  # geom_text_repel(data=subset(Vax_TSA_now, Vax_pct < 0.38 | 
  #                         Vax_pct > 0.65 |
  #                         Blueness > 0.55),
  ggrepel::geom_text_repel(
            aes(Blueness,Excess_percap,label=TSA), hjust=0,
                nudge_x=.005) +
  labs(title="Excess Deaths (per 100,000) vs. Biden percent vote",
       subtitle="By Trauma Service Area",
       x="Biden Percent Vote",
       y="Excess Deaths per 100,000")


#  By TSA, COVID deaths

By_TSA %>% 
  mutate(Covid_percap=Covid/Population*100000) %>% 
  ggplot(aes(x=Blueness,y=Covid_percap ))+
  geom_point()+
  geom_smooth(method=lm)+
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  ggrepel::geom_text_repel(
            aes(Blueness,Covid_percap,label=TSA), hjust=0,
                nudge_x=.005) +
  labs(title="Covid Deaths (per 100,000) vs. Biden percent vote",
       subtitle="By Trauma Service Area",
       x="Biden Percent Vote",
       y="Covid Deaths per 100,000")

#  Trajectory from Excell to Covid

By_TSA <- By_TSA %>% 
  mutate(Covid_percap=Covid/Population*100000) %>% 
  mutate(Excess_percap=Excess/Population*100000)

Labels <- By_TSA %>% 
  select(TSA, Blueness, Covid_percap)
  
TSA_longer <- By_TSA %>% 
  select(TSA, Blueness, Covid_percap, Excess_percap) %>% 
  pivot_longer(!c(TSA, Blueness), names_to="Names", values_to="Values")

#By_TSA %>% 
#  select(TSA, Blueness, Covid_percap, Excess_percap) %>% 
#  pivot_longer(!c(TSA, Blueness), names_to="Names", values_to="Values") %>% 
p <-   ggplot(data=TSA_longer, aes(x=Blueness,y=Values, color=Names ))+
  geom_point()+
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  geom_segment(data = By_TSA, 
               aes(x = Blueness, xend = Blueness, 
                   y = Covid_percap,  yend = Excess_percap, 
                   group = TSA), 
               colour = "black", 
               arrow = arrow(length=unit(0.2, "cm"))) +
   ggrepel::geom_text_repel(data=subset(TSA_longer, Names=="Covid_percap"),
             mapping=aes(x=Blueness,y=Values,label=TSA), hjust=0,
                 nudge_x=.005) +
  labs(title="Covid and Excess Deaths (per 100,000) vs. Biden percent vote",
       subtitle="By Trauma Service Area",
       x="Biden Percent Vote",
       y="Covid or Excess Deaths per 100,000")



```





