---
title: Covid Vaccine
author: Alan Jackson
date: '2021-01-03'
slug: [covid-vaccine]
categories:
  - Covid-19
  - Data Fitting
tags:
  - Covid-19
  - Texas
keywords:
  - tech
---

```{r setup, include=FALSE}

library("tidyverse")
library("stringr") # string tools
library("lubridate")

path <-  "/home/ajackson/Dropbox/Rprojects/Covid/"

df <- readRDS(paste0(path, "Vaccinations.rds"))
covid <- readRDS(paste0(path, "Covid.rds"))

knitr::opts_chunk$set(echo=TRUE)
```

December 2020 vaccine distribution and administration began. I started trapping 
the daily spreadsheet from the state health department that tracked progress.
This blog entry is really for prototyping some of the data cleanup and displays
that I will incorporate into my shiny app.

Let's take a look at data issues.

```{r}

df %>% 
  filter(!is.na(Pct_given)) %>% 
  ggplot(aes(x=Pct_given)) +
  geom_histogram()+
  labs(x="Percent Distributed Administered",
       title="Distribution of Administered Vaccine")
  
#   after cleaning

df %>% 
  filter(!is.na(Pct_given)) %>% 
  filter(Pct_given<100) %>% 
  ggplot(aes(x=Pct_given)) +
  geom_histogram(binwidth=2)+
  labs(x="Percent Distributed Administered",
       title="Distribution of Administered Vaccine")
```

Well, that's irritating. Some folks are giving shots they don't have.

```{r}

#   Negative doses
df %>% 
  filter(County != "Texas") %>% 
  group_by(County) %>% 
    mutate(min_admin=min(Daily_adm)) %>% 
  ungroup() %>% 
  arrange(County) %>% 
  filter(min_admin<0) %>% 
  ggplot(aes(x=Date, y=Doses_admin, color=County)) +
  geom_line()+
  theme(legend.position = "none") + 
  labs(x="Date", y="Cumulative Doses",
       title="Administered Vaccine")
```

So we have a few counties where the cumulative numbers have decreased. That is
disappointing.

```{r}

# how fair is the distribution?

df %>% 
  group_by(County) %>% 
    mutate(fairness=10000*Doses_alloc/Pop_adult) %>% 
    summarise(fairness=max(fairness, na.rm=TRUE)) %>% 
  ggplot(aes(x=fairness)) +
  geom_histogram()+
  labs(x="Per Capita Doses Distributed",
       title="Distribution of Administered Vaccine")

df %>% 
  filter(County!="Texas") %>% 
  group_by(County) %>% 
    mutate(fairness=10000*Doses_alloc/Pop_adult) %>% 
    summarise(fairness=max(fairness, na.rm=TRUE), Pop_adult=max(Pop_adult, na.rm=TRUE)) %>% 
  ggplot(aes(x=fairness, y=Pop_adult)) + 
  geom_point() + 
  labs(x="Per Capita Doses Distributed (Doses per 10,000)",
       y="County Population",
       title="Distribution of Administered Vaccine")


df %>% 
  filter(County!="Texas") %>% 
  group_by(County) %>% 
    summarise(Doses_alloc=max(Doses_alloc, na.rm=TRUE), Pop_adult=max(Pop_adult, na.rm=TRUE)) %>% 
  ggplot(aes(x=Doses_alloc, y=Pop_adult)) + 
  geom_point() + 
  scale_x_log10() + 
  scale_y_log10() +
  labs(x="Log Doses Distributed",
       y="Log County Population",
       title="Distribution of Administered Vaccine")
```

Looks pretty fair. At low numbers can see the effect of minimum vaccine lot size.

```{r}

#   % over time

df %>% 
  group_by(County) %>% 
    mutate(Pct_Coverage=100*Doses_admin/Pop_adult) %>% 
  ungroup() %>% 
  filter(Pct_Coverage<25) %>% 
  ggplot(aes(x=Date, y=Pct_Coverage, color=County)) + 
  geom_line() + 
  theme(legend.position = "none") + 
  labs(x="Date",
       y="Percent Coverage",
       title="Percent of each County Vaccinated")  
  
```

Cottle county jumps to almost 140%, the rest are below 25%.

```{r}

#   How to best calculate the date when everyone is vaccinated?

#   Calculate an average inoculation rate over the last 7 days

this_day <- lubridate::today()

df %>% 
  filter(Date>last(Date)-7) %>% 
  group_by(County) %>% 
    summarize(Inoculation_rate=mean(Daily_adm, na.rm=TRUE),
              Pop_adult=last(Pop_adult),
              People_one_dose=last(People_one_dose)) %>% 
  mutate(Finish_date=as_date(this_day+((Pop_adult-People_one_dose)/Inoculation_rate+28))) %>% 
  mutate(DaysFromNow=((Pop_adult-People_one_dose)/Inoculation_rate+28)) %>% 
  ggplot(aes(x=DaysFromNow/7)) +
  geom_histogram() +
  labs(x="Weeks from Today",
       title="Weeks from present to reach full vaccination by county",
       subtitle="Rate derived from last 7 days")

```






















