---
title: Covid Testing In Texas
author: Alan Jackson
date: '2020-07-05'
slug: covid-testing-in-texas.en-us
categories:
  - Covid-19
tags:
  - Covid-19
  - Texas
keywords:
  - tech
---

## Covid tests in Texas

The second entry to look at the testing in Texas. This time there is much
more data to examine, but also a new and irritating problem. At some point, 
for some counties, antibody testing got mixed in with the PCR tests, so the
numbers are not nearly as good as they should be. Last I read, about 10% of
the tests are the wrong test, but I suspect this is not evenly distributed
by county, but rather concentrated in a few. The total test numbers for the
state are supposed to be free of this issue, so I'll also look at those.

The raw data comes from the [State Health Department website](https://txdshs.maps.arcgis.com/apps/opsdashboard/index.html#/ed483ecd702b4298ab01e8b9cafc8b83).

```{r setup, include=FALSE}

library(tidyverse)
library(ggplot2)
library(lubridate)
library(purrr)
library(broom)

path <- "/home/ajackson/Dropbox/Rprojects/Covid/Today_Data/"
Tests_Total <- readRDS(paste0(path, "Today_TestingData.rds"))

Tests_Total <- Tests_Total %>% 
  mutate(Total=as.numeric(Total))

Tests_by_county <- readRDS(paste0(path, "Today_County_Testing.rds"))

Tests_by_county <- Tests_by_county %>% 
  mutate(Date=ymd(Date))

Covid_data <- readRDS("/home/ajackson/Dropbox/Rprojects/Covid/Covid.rds")

Covid_data <- Covid_data %>% 
  mutate(Days=as.integer(Date-ymd("2020-03-10")))
```

## Compare totals to total

Let's look at differences between the state total and the sum of the
county numbers.

```{r compare}

county <- Tests_by_county %>% 
  filter(!County=="TOTAL") %>% 
  group_by(Date) %>% 
    summarise(County=sum(Tests))

foo <- Tests_by_county %>% filter(County=="TOTAL")

left_join(Tests_Total, county, by="Date") %>% 
  select(Date, State=Total, County) %>% 
  pivot_longer(-Date,  names_to="Source", values_to="Test_Total") %>% 
ggplot(aes(x=Date, y=Test_Total, color=Source)) + 
  geom_line() + 
  geom_line(data=foo, aes(x=Date, y=Tests), color="red")+
  labs(title="Number of Tests, Texas", y="Total Tests")

```

So good news, the figures have been updated, hopefully to include only
relevant tests and not the antibody tests.

##  Positive tests

So let's join the tests dataset with the positive cases dataset and get
a measure of percent positive

```{r pct positive}

full_data <- left_join(Covid_data, 
                      Tests_by_county, 
                      by=c("Date", "County")) %>% 
  filter(!grepl("Probable.*", County))

full_data <- full_data %>% 
  group_by(Date) %>% 
    mutate(Test_Total=sum(Tests, na.rm=TRUE), Case_Total=sum(Cases, na.rm=TRUE)) %>% 
  mutate(pct_pos=Cases/Tests*100)

# Plot total

full_data %>% 
  filter(Test_Total>0) %>% 
  group_by(Date) %>% 
    summarise(Test_Total=sum(Tests, na.rm=TRUE), 
              Case_Total=sum(Cases, na.rm=TRUE)) %>%
  mutate(pct_pos=Case_Total/Test_Total*100) %>% 
  ggplot(aes(x=Date, y=pct_pos)) +
    geom_line() + 
    labs(title="Percent Positive Tests for Texas", x="Date", y="Percent")

```

Not good. The percentage is rising rapidly. Just for clarification, let's
calculate the daily tests and cases instead of looking at integrated values.

```{r dailies}

full_data <- full_data %>% 
  group_by(County) %>% 
  mutate(Daily_tests=Tests-lag(Tests, default=0),
         Daily_Cases=Cases-lag(Cases, default=0))

#   and let's look at it

foo <- 
full_data %>% 
  filter(Daily_tests>0) %>% 
  group_by(Date) %>% 
    summarise(Daily_Test_Total=sum(Daily_tests, na.rm=TRUE), 
              Daily_Case_Total=sum(Daily_Cases, na.rm=TRUE))

foo %>% 
  ggplot(aes(x=Date, y=Daily_Test_Total)) + 
  geom_point() +
  geom_smooth() + 
  theme(legend.position = "none") +
  labs(title="Texas Daily Test Total", y="Daily Tests")

foo %>% 
  mutate(Daily_pct=Daily_Case_Total/Daily_Test_Total*100) %>%
  filter(Daily_pct<50) %>% 
  ggplot(aes(x=Date, y=Daily_pct)) + 
  geom_point() +
  geom_smooth() + 
  theme(legend.position = "none") +
  labs(title="Texas Daily Positive Test Percent", y="Percent")

```

Definitely looks like things going in the wrong direction. But the data is
also surprisingly noisy. Let's spend some time looking into the data
quality.

```{r quality}

#  Look for large deviations in daily tests

foo <- full_data %>% 
  filter(Daily_tests>0) %>% 
  mutate(pct=Daily_Cases/Daily_tests*100) %>% 
  filter(pct>20) %>% 
  select(County, Date, Cases, Daily_tests, Daily_Cases, pct)
foo %>% 
    ggplot(aes(x=Date, y=pct)) + 
    geom_point()

#   Okay let's look at > 2000 first

foo %>% filter(pct>2000) %>% 
    ggplot(aes(x=Date, y=pct, color=County)) + 
    geom_point()

# Let's plot full dataset for all counties with pct > 2000

foo <- full_data %>% 
  filter(Daily_tests>0) %>% 
  mutate(pct=Daily_Cases/Daily_tests*100) %>% 
  group_by(County) %>% 
    mutate(maxpct=max(pct, na.rm = TRUE)) %>% 
  ungroup() %>% 
  select(County, Date, Cases, Daily_tests, Daily_Cases, pct, maxpct)

foo %>% filter(maxpct>2000) %>% 
    ggplot(aes(x=Date, y=pct, color=County)) + 
      geom_point() + 
      geom_line()

foo %>% filter(maxpct>2000) %>% 
    ggplot(aes(x=Date, y=Daily_tests, color=County)) + 
      geom_point() + 
      geom_line()


```

Wow. Data is super noisy. Clearly some pre-processing is in order to make
the data more usable. And those annoying gaps. Those are in the original
data.

Let's look at some individual counties

```{r singles}

foo <- full_data %>% 
  filter(Daily_tests>0) %>% 
  mutate(pct=Daily_Cases/Daily_tests*100) %>% 
  group_by(County) %>% 
    mutate(maxpct=max(pct, na.rm = TRUE)) %>% 
  ungroup() %>% 
  select(County, Date, Cases, Daily_tests, Daily_Cases, pct, maxpct)

foo %>% 
  filter(County=="Bexar") %>% 
    ggplot(aes(x=Date, y=pct)) + 
      geom_point() + 
      geom_line()

foo %>% 
  filter(County=="Bexar") %>% 
    ggplot(aes(x=Date, y=Daily_tests)) + 
      geom_point() + 
      geom_line()

#   and how about trying some rolling averages

bexar <- foo %>% filter(County=="Bexar")

vars <- c("avg3", "avg5", "avg7", "avg9")

window <-  c(3,5,7,9)
for (i in 1:4) {
  bexar <- bexar %>% 
    mutate(!!vars[i] := zoo::rollapply(Daily_tests, window[i], 
                  FUN=function(x) mean(x, na.rm=TRUE),
                  fill=c(first(Daily_tests), NA, last(Daily_tests))))
}

bexar %>% select(-County, -Daily_Cases,-Cases, -pct, -maxpct) %>%
  pivot_longer(c(-Date, -Daily_tests), 
               names_to="Length",
               values_to="Averages") %>% 
    ggplot(aes(x=Date, y=Daily_tests)) + 
      geom_point() + 
      geom_line(aes(y=Averages, color=Length)) +
      geom_smooth(aes(y=Daily_tests)) +
      labs(title="Bexar County Daily Tests Averaged Over Various Windows",
           y="Daily Tests")
```

Even a short 3-day window does a pretty good job of cleaning up the
jitter in the data - for the most part all the windows give very similar
answers. Arbitrarily I will pick the 5-point window as enough. Since we 
want to look at ratios of tests to cases, let's also look at the daily
case numbers with a similar analysis.

```{r case daily}

foo <- full_data %>% 
  group_by(County) %>% 
    mutate(maxcase=max(Cases, na.rm = TRUE)) %>% 
  ungroup() %>% 
  select(County, Date, Cases, Daily_tests, Daily_Cases, maxcase) %>% 
  filter(maxcase>=30)

foo %>% 
  ggplot(aes(x=Date, y=Daily_Cases, color=County)) + 
  geom_line() +
  theme(legend.position = "none")

```

Daily case numbers are pretty noisy as well. To usefully do anything with
them and the test numbers, smoothing will be required.

Let's try it out for Harris county, since that has suitably large
deviations.

```{r}
harris <- foo %>% filter(County=="Harris")

harris <- harris %>% 
  filter(Daily_tests>0) %>% 
  mutate(avg_tests := zoo::rollapply(Daily_tests, 5, 
                FUN=function(x) mean(x, na.rm=TRUE),
                fill=c(first(Daily_tests), NA, last(Daily_tests)))) %>% 
  mutate(avg_cases := zoo::rollapply(Daily_Cases, 5, 
                FUN=function(x) mean(x, na.rm=TRUE),
                fill=c(first(Daily_Cases), NA, last(Daily_Cases)))) %>% 
  mutate(pct=avg_cases/avg_tests*100, 
         rawpct=Daily_Cases/Daily_tests*100 )

harris %>% 
  select(Date, pct, rawpct) %>% 
  pivot_longer(-Date, names_to="Smoothing", values_to="Percent") %>% 
    ggplot(aes(x=Date, y=Percent)) + 
      geom_point() + 
      geom_line() +
      geom_smooth() +
      facet_wrap(~ Smoothing, scales = "free_y") +
      labs(title="Harris County Daily Cases/Tests",
           y="Percent")
```

Truthfully, I'm liking the loess fit more and more, as it really blows
through what are obviously bad data areas. What if we use a loess
prior to calculating the percent?

```{r loess trials}

models <- foo %>%
  mutate(Days=as.integer(Date-ymd("2020-03-10"))) %>% 
  select(County, Days, Daily_tests, Daily_Cases) %>% 
  mutate(rawpct=Daily_Cases/Daily_tests*100) %>% 
  filter(!is.na(Daily_tests)) %>% 
  filter(Daily_tests>0) %>% 
        tidyr::nest(-County) %>%
        dplyr::mutate(
                # Perform loess calculation on each County group
                m_case = purrr::map(data, loess,
                               formula = Daily_Cases ~ Days, span = .5),
                # Retrieve the fitted values from each model
                fitcase = purrr::map(m_case, `[[`, "fitted"),
                m_test = purrr::map(data, loess,
                               formula = Daily_tests ~ Days, span = .5),
                # Retrieve the fitted values from each model
                fittest = purrr::map(m_test, `[[`, "fitted"),
                m_pct = purrr::map(data, loess,
                               formula = rawpct ~ Days, span = .5),
                # Retrieve the fitted values from each model
                fitpct = purrr::map(m_pct, `[[`, "fitted")
        )

# Apply fitted y's as a new column
results <- models %>%
        dplyr::select(-m_case, -m_test, -m_pct) %>%
        tidyr::unnest(c(data, fitcase, fittest, fitpct)) %>% 
        mutate(Date=ymd("2020-03-10")+Days) %>% 
        mutate(pct=fitcase/fittest*100)

# Plot with loess line for each group
ggplot(results, aes(x = Date, y = pct, group = County, colour = County)) +
        geom_point() +
        geom_line() +
    theme(legend.position = "none")


```

Hmmm... issues. Let's start looking at the individual counties that have
the biggest issues.

```{r county pct issues}

topten <- results %>% 
  arrange(desc(pct)) %>% 
  select(County) %>% 
  unique() %>% 
  head(10)

results %>% 
  filter(County %in% topten[[1]]) %>% 
  select(County, Date, pct, rawpct, fitpct) %>% 
  pivot_longer(c(-Date, -County), names_to="Algorithm", values_to="Percent") %>% 
    ggplot(aes(x=Date, y=Percent, color=Algorithm)) + 
      geom_point() + 
      geom_line() +
      facet_wrap(~ County, scales = "free_y") +
      labs(title="County Daily Cases/Tests",
           y="Percent")

results %>% 
  filter(County %in% topten[[1]]) %>% 
  select(County, Date, Daily_Cases, Daily_tests) %>% 
  pivot_longer(c(-Date, -County), names_to="Tests_Cases", values_to="Number") %>% 
    ggplot(aes(x=Date, y=Number, color=Tests_Cases)) + 
      geom_point() + 
      geom_line() +
      facet_wrap(~ County, scales = "free_y") +
      labs(title="County Daily Cases and Tests",
           y="Number")
```

Wow. Kind of a mess. Houston county I suspect someone got confused with city of Houston. Moore county is probably real - meat-packing or a prison - Jones looks suspicious. Bee county is a prison?

Let's smooth everything with a loess on the cumulative numbers. First we'll
do it for a few counties to test the loess parameters.

```{r loess on cums}

counties <- c("Harris", "Dallas", "Tarrant", "Bexar", "Travis",
              "Collin", "Denton", "Hidalgo", "El Paso", "Fort Bend",
              "Williamson")

counties <- c("Bexar", "Tarrant",  "Williamson")
models <- full_data %>%
  filter(County %in% counties) %>% 
  filter(Tests>0) %>% 
  mutate(Days=as.integer(Date-ymd("2020-03-10"))) %>% 
  select(County, Days, Tests, Cases) %>% 
        tidyr::nest(-County) %>%
        dplyr::mutate(
                # Perform loess calculation on each County group
                m_case = purrr::map(data, loess,
                               formula = Cases ~ Days, span = .5),
                # Retrieve the fitted values from each model
                fitcase = purrr::map(m_case, `[[`, "fitted"),
                m_test = purrr::map(data, loess,
                               formula = Tests ~ Days, span = .5),
                # Retrieve the fitted values from each model
                fittest = purrr::map(m_test, `[[`, "fitted"),
        )

# Apply fitted y's as a new column
results <- models %>%
  dplyr::select(-m_case, -m_test) %>%
  tidyr::unnest(c(data, fitcase, fittest)) %>% 
  mutate(Date=ymd("2020-03-10")+Days) %>% 
  group_by(County) %>% 
    mutate(Daily_tests=fittest-dplyr::lag(fittest, default=NA),
           Daily_Cases=fitcase-dplyr::lag(fitcase, default=NA)) %>% 
    mutate(pct=Daily_Cases/Daily_tests*100) %>% 
 ungroup()
  

# Plot with loess line for each group
ggplot(results, aes(x = Date, y = fitcase, group = County, colour = County)) +
        geom_point() +
        geom_point(aes(y=fittest))

#   Williamson and Bexar counties
    
full_data %>% 
  filter(County=="Williamson") %>% 
  ggplot(aes(x=Date)) +
    geom_smooth(aes(y=Cases)) +
    geom_point(aes(y=Cases)) +
    geom_smooth(aes(y=Tests)) +
    geom_point(aes(y=Tests)) 
  
results %>% 
  filter(County=="Williamson") %>% 
  ggplot(aes(x=Date)) +
    geom_smooth(aes(y=Daily_Cases)) +
    geom_point(aes(y=Daily_Cases)) +
    geom_smooth(aes(y=Daily_tests)) +
    geom_point(aes(y=Daily_tests)) 

```

Hmmm.... I need to fill in the holes that cause the derivative to blow up.
A simple linear interpolation ought to be just fine.

```{r linear interpolate to fill holes}

#   First trim NA's from in front of data, then interpolate gaps

foo <- full_data %>% 
  filter(Days>41) %>%
  filter(Days<116) %>%
  group_by(County) %>%
    mutate(Tests = zoo::na.approx(Tests, Days, na.rm=FALSE)) %>% 
  ungroup()


counties <- c("Harris", "Dallas", "Tarrant", "Bexar", "Travis",
              "Collin", "Denton", "Hidalgo", "El Paso", "Fort Bend",
              "Williamson")
counties <- c("Bexar",   "Williamson")

models <- foo %>%
  filter(County %in% counties) %>% 
  mutate(Days=as.integer(Date-ymd("2020-03-10"))) %>% 
  select(County, Days, Tests, Cases) %>% 
        tidyr::nest(-County) %>%
        dplyr::mutate(
                # Perform loess calculation on each County group
                m_case = purrr::map(data, loess,
                               formula = Cases ~ Days, span = .5),
                # Retrieve the fitted values from each model
                fitcase = purrr::map(m_case, `[[`, "fitted"),
                m_test = purrr::map(data, loess,
                               formula = Tests ~ Days, span = .5),
                # Retrieve the fitted values from each model
                fittest = purrr::map(m_test, `[[`, "fitted"),
        )

# Apply fitted y's as a new column
results <- models %>%
  dplyr::select(-m_case, -m_test) %>%
  tidyr::unnest(c(data, fitcase, fittest)) %>% 
  mutate(Date=ymd("2020-03-10")+Days) %>% 
  filter(!is.na(Tests)) %>% 
  group_by(County) %>% 
    mutate(Daily_tests=fittest-dplyr::lag(fittest, default=NA),
           Daily_Cases=fitcase-dplyr::lag(fitcase, default=NA)) %>% 
    mutate(pct=Daily_Cases/Daily_tests*100) %>% 
 ungroup()
  

# Plot with loess line for each group
ggplot(results, aes(x = Date, y = fitcase, group = County, colour = County)) +
        geom_point() +
        geom_point(aes(y=fittest))

# Now let's try various span values

for (span in (1:5)*0.1) {
  
models <- foo %>%
  filter(County %in% counties) %>% 
  mutate(Days=as.integer(Date-ymd("2020-03-10"))) %>% 
  select(County, Days, Tests, Cases) %>% 
        tidyr::nest(-County) %>%
        dplyr::mutate(
                # Perform loess calculation on each County group
                m_case = purrr::map(data, loess,
                               formula = Cases ~ Days, span = span),
                # Retrieve the fitted values from each model
                fitcase = purrr::map(m_case, `[[`, "fitted"),
                m_test = purrr::map(data, loess,
                               formula = Tests ~ Days, span = span),
                # Retrieve the fitted values from each model
                fittest = purrr::map(m_test, `[[`, "fitted"),
        )

# Apply fitted y's as a new column
results <- models %>%
  dplyr::select(-m_case, -m_test) %>%
  tidyr::unnest(c(data, fitcase, fittest)) %>% 
  mutate(Date=ymd("2020-03-10")+Days) %>% 
  filter(!is.na(Tests)) %>% 
  group_by(County) %>% 
    mutate(Daily_tests=fittest-dplyr::lag(fittest, default=NA),
           Daily_Cases=fitcase-dplyr::lag(fitcase, default=NA)) %>% 
    mutate(pct=Daily_Cases/Daily_tests*100) %>% 
 ungroup()
  

# Plot with loess line for each group
p <- 
ggplot(results, aes(x = Date, y = fitcase, group = County, colour = County)) +
        geom_point() +
        geom_point(aes(y=fittest)) +
        labs(title=paste("Span =", span))

print(p)

p <- 
ggplot(results, aes(x = Date, y = Daily_tests, group = County, colour = County)) +
        geom_point() +
        labs(title=paste("Span =", span))

print(p)
  
p <- 
ggplot(results, aes(x = Date, y = pct, group = County, colour = County)) +
        geom_point() +
        labs(title=paste("Span =", span))

print(p)
}


```

New strategy. Let's find all the counties where the daily tests go negative,
and look closely at those for a potential correction factor, county by 
county.

```{r find bad counties}
#   First trim NA's from in front of data, then interpolate gaps

foo <- full_data %>% 
  filter(Days>41) %>%
  filter(Days<116) %>%
  group_by(County) %>%
    mutate(Tests = zoo::na.approx(Tests, Days, na.rm=FALSE)) %>% 
    mutate(Daily_tests=Tests-lag(Tests, default=0),
         Daily_Cases=Cases-lag(Cases, default=0)) %>% 
  ungroup()

foo %>% filter(Daily_tests<0,
               Daily_tests>-500) %>% 
  ggplot(aes(x=Daily_tests))+
    geom_histogram()

# Ignore small fry negatives, let's look at daily tests <-100 
#  and total > 500

badcounties <- foo %>% 
  filter(Daily_tests<=-200) %>% 
  filter(Tests>500) %>% 
  select(County) %>% 
  unique()

badcounties <- badcounties[[1]]

foo %>% filter(County %in% badcounties) %>% 
  ggplot(aes(x=Date, y=Daily_tests, color=County))+
  geom_point()
    theme(legend.position = "none")
    
foo %>% filter(County %in% badcounties) %>% 
  ggplot(aes(x=Date, y=Tests, color=County))+
      geom_point() + 
      geom_line() +
      facet_wrap(~ County, scales = "free_y") +
      labs(title="County Tests",
           y="Number")

bexar <- foo %>% filter(County=="Bexar") 
  #  delete from 2020-05-11 to 2020-06-03
bexar$Tests[between(bexar$Date, 
                    as.Date("2020-05-11"), 
                    as.Date("2020-06-03"))] <- NA
  #  reinterpolate
bexar <- bexar %>% 
  mutate(Tests = zoo::na.approx(Tests, Days, na.rm=FALSE))

bexar %>% 
  ggplot(aes(x=Date, y=Tests))+
      geom_point() + 
      geom_line() +
      labs(title="Bexar County Tests",
           y="Number")
```

Basically the test data prior to June 4 is crap. With much effort I could 
fake in new data, or I could just ignore data prior to June 4. Let's go with
plan B.

```{r Late data only}

#   Clean up the data (gaps), smooth cases and tests with a long loess,
#   calculate daily numbers, and then the ratio

#   First trim NA's from in front of data, then interpolate gaps

foo <- full_data %>% 
  filter(Date>as.Date("2020-06-03")) %>%
  filter(Days<116) %>%
  group_by(County) %>%
    mutate(Tests = zoo::na.approx(Tests, Days, na.rm=FALSE)) %>% 
    mutate(maxcases=max(Cases)) %>% 
  ungroup() %>% 
  filter(maxcases>=30)



models <- foo %>%
  mutate(Days=as.integer(Date-ymd("2020-03-10"))) %>% 
  select(County, Days, Tests, Cases) %>% 
        tidyr::nest(-County) %>%
        dplyr::mutate(
                # Perform loess calculation on each County group
                m_case = purrr::map(data, loess,
                               formula = Cases ~ Days, span = .7),
                # Retrieve the fitted values from each model
                fitcase = purrr::map(m_case, `[[`, "fitted"),
                m_test = purrr::map(data, loess,
                               formula = Tests ~ Days, span = .7),
                # Retrieve the fitted values from each model
                fittest = purrr::map(m_test, `[[`, "fitted"),
        )

# Apply fitted y's as a new column
results <- models %>%
  dplyr::select(-m_case, -m_test) %>%
  tidyr::unnest(c(data, fitcase, fittest)) %>% 
  mutate(Date=ymd("2020-03-10")+Days) %>% 
  filter(!is.na(Tests)) %>% 
  group_by(County) %>% 
    mutate(Daily_tests=fittest-dplyr::lag(fittest, default=NA),
           Daily_Cases=fitcase-dplyr::lag(fitcase, default=NA)) %>% 
    mutate(pct=Daily_Cases/Daily_tests*100) %>% 
 ungroup()
  
County_list <- unique(results$County)


# Plot with loess line for each group
for (i in seq(18,162,18) ) {
  p <- results %>% filter(County %in% County_list[(i-17):i]) %>% 
  ggplot(aes(x = Date, y = Daily_tests, group = County, colour = County)) +
    geom_line() +
    geom_line(aes(y=Daily_Cases), color="black")+
    theme(legend.position = "none")+
    facet_wrap(~ County, scales = "free_y")
  
  print(p)
}
```

That looks pretty good. About as good as it is going to get. All I need
to do now is eliminate the zero and negative tests and cases, and set 
some cutoffs for cases/tests.

So let's start it all from scratch, so I'll have some tested, portable
code at the end of the day.

Also try out several plotting schemes.

```{r final}

path <- "/home/ajackson/Dropbox/Rprojects/Covid/Today_Data/"

Tests_by_county <- readRDS(paste0(path, "Today_County_Testing.rds"))

Tests_by_county <- Tests_by_county %>% 
  mutate(Date=ymd(Date))

Covid_data <- readRDS("/home/ajackson/Dropbox/Rprojects/Covid/Covid.rds")

Covid_data <- Covid_data %>% 
  mutate(Days=as.integer(Date-ymd("2020-03-10")))

full_data <- left_join(Covid_data, 
                      Tests_by_county, 
                      by=c("Date", "County")) %>% 
  filter(!grepl("Probable.*", County))

#   De-step

Prison_counties <- c("Jones", "Anderson", "Walker", "Medina", "Rusk",
                     "Grimes", "Coryell", "Houston", "Pecos", "Angelina",
                     "Bowie", "Jefferson", "Brazoria")

full_data <- full_data %>% 
  mutate(Raw_cases=Cases) %>% 
  arrange(County) %>% 
  group_by(County) %>% 
    mutate(delta=Cases-lag(Cases)) %>% 
    replace_na(list(delta=0)) %>% 
    mutate(Threshold=as.numeric((abs(delta)>25)&(abs(delta)/Cases>0.10))*delta) %>% 
    mutate(Threshold=cumsum(Threshold)) %>%  
  ungroup() %>% 
  mutate(Cases=ifelse(County %in% Prison_counties, 
                      Raw_cases-Threshold, 
                      Raw_cases)) %>% 
  select(-Threshold, -delta)

#full_data <- full_data %>% 
#  group_by(Date) %>% 
#    mutate(Test_Total=sum(Tests, na.rm=TRUE), Case_Total=sum(Cases, na.rm=TRUE)) %>% 
#  mutate(pct_pos=Cases/Tests*100)
#
#   First trim crap prior to June 4, then interpolate gaps

foo <- full_data %>% 
  filter(Date>as.Date("2020-06-03")) %>%
  group_by(County) %>%
    mutate(Tests = zoo::na.approx(Tests, Days, na.rm=FALSE)) %>% 
    mutate(maxcases=max(Cases)) %>% 
  ungroup() %>% 
  filter(maxcases>=500) %>% 
  filter(!is.na(Tests))



models <- foo %>%
  mutate(Days=as.integer(Date-ymd("2020-03-10"))) %>% 
  select(County, Days, Tests, Cases) %>% 
        tidyr::nest(-County) %>%
        dplyr::mutate(
                # Perform loess calculation on each County group
                m_case = purrr::map(data, loess,
                               formula = Cases ~ Days, span = .7),
                # Retrieve the fitted values from each model
                fitcase = purrr::map(m_case, `[[`, "fitted"),
                m_test = purrr::map(data, loess,
                               formula = Tests ~ Days, span = .7),
                # Retrieve the fitted values from each model
                fittest = purrr::map(m_test, `[[`, "fitted"),
        )

# Apply fitted y's as a new column
results <- models %>%
  dplyr::select(-m_case, -m_test) %>%
  tidyr::unnest(c(data, fitcase, fittest)) %>% 
  mutate(Date=ymd("2020-03-10")+Days) %>% 
  filter(!is.na(Tests)) %>% 
  group_by(County) %>% 
    mutate(Daily_tests=fittest-dplyr::lag(fittest, default=0),
           Daily_Cases=fitcase-dplyr::lag(fitcase, default=0)) %>% 
    mutate(pct=Daily_Cases/Daily_tests*100) %>% 
 ungroup()

# Now smooth the daily values

case_loess <- function(){
  
}

models2 <- results %>%
  select(County, Days, Daily_tests, Daily_Cases) %>% 
        tidyr::nest(-County) %>%
        dplyr::mutate(
                # Perform loess calculation on each County group
                m_case = purrr::map(data, loess,
                               formula = Daily_Cases ~ Days, span = .7,
                               na.action = na.exclude),
                # Retrieve the fitted values from each model
                fitcase = purrr::map(m_case, `[[`, "fitted"),
                
                m_test = purrr::map(data, loess,
                               formula = Daily_tests ~ Days, span = .7,
                               na.action = na.exclude),
                # Retrieve the fitted values from each model
                fittest = purrr::map(m_test, `[[`, "fitted"),
        ) 

# Apply fitted y's as a new column
results2 <- models2 %>%
  dplyr::select(-m_case, -m_test) %>%
  tidyr::unnest(c(data, fitcase, fittest)) %>% 
  mutate(Date=ymd("2020-03-10")+Days) %>% 
  group_by(County) %>% 
    mutate(pct=fitcase/fittest*100) %>% 
 ungroup()


  
County_list <- unique(results$County)

results2 %>%
  mutate(pct=ifelse(pct>100, NA, pct)) %>% 
  mutate(pct=ifelse(pct<0,   NA, pct)) %>% 
  group_by(County) %>% 
    mutate(maxpct=max(pct, na.rm=TRUE)) %>% 
  ungroup() %>% 
 # filter(maxpct>100) %>% 
  #ggplot(aes(x=Date, y=Daily_tests)) + 
  ggplot(aes(x=Date, y=pct)) + 
  #theme(legend.position = "none", text = element_text(size=20)) +
  geom_line(aes(group=County, color=County)) 
  #geom_line(aes(group=County, color=County)) +
  #geom_point(aes(y=Daily_tests, color=County))

```

I give up. The data is so badly compromised that I can't see making anything useful out of it.











