---
title: City Permits Online
author: Alan Jackson
date: '2020-01-02'
slug: city-permits-online
categories:
  - Infrastructure
  - Permits
tags:
  - Houston
  - Mapping
keywords:
  - tech
---

```{r setup, include=FALSE}

library("tidyverse")
library("stringr") # string tools
library("rvest")# web scraping
library("lettercase") # fiddle with letter case
library("lubridate")
library(sf)

knitr::opts_chunk$set(echo=TRUE)
```

The city of Houston makes a file available on the web every week containing a summary of the past week's building permits. I found this file a bit difficult to digest - it needed a map, it needed search and filtering. So I wrote some code to automatically read the file in each week, merge it with the previous weeks files, and then upload that to the web where I have an application to display it. Oh, I also geocoded the addresses in the file, so they could be mapped.

The viewer is of course a [shiny app.](https://ajackson.shinyapps.io/ViewPermitData/)

## Gather the data

The data is found at http://houstontx.gov/planning/Publications/listsrv/WebeReport.html

```{r read in the data}

options(stringsAsFactors = FALSE)

Weekly_Rpt_URL <- "http://houstontx.gov/planning/Publications/listsrv/WebeReport.html"

## Read in a single weekly report

rpt <- Weekly_Rpt_URL %>% 
  read_html() %>% 
  html_nodes(css="table") %>%
  html_table(header=FALSE, fill=TRUE)
  
rpt <- rpt[[1]]

```

It is contained in an html table, so that is pretty easy to read in. 

In the header for the table it has the start and end dates for the table - let's pull those out, and then delete the bits of the table we don't want. Then we can rename the columns to be useful.

```{r clean up table}
#   First pull start and end dates out

Start_Date <- strsplit(rpt$X1[3],"\n")[[1]][2]
End_Date <- strsplit(rpt$X1[4],"\n")[[1]][2]

#   delete unwanted columns and rows

rpt <- rpt %>% 
  select(-c(X7, X8, X9, X10)) %>%  # drop bad columns
  slice(14:n()-7) #  drop first 7 and last 7 rows
  

#   rename columns

names(rpt) <- c("Zip", "Permit_Date", "Permit_Type", "Permit_Number", "Address", "Comments")


```

## Geocoding

In order to map the data, we need to attach lat long coordinates to each point. Currently, to do that, I am using the Census Bureau server to geocode the addresses. It is far from perfect, but see my previous post on geocoding for details.

Note that one minor (but irritating) issue is that for some addresses the census geocoder changes the city Houston to "Clutch City". So I have to deal with that.

```{r Geocoding, results="hide"}


census_geocoder <- function(address,type,secondary,state){

    library(jsonlite)
    library(RCurl)

    addy <- paste("street=",gsub(" ","+",address),sep="")
    if(type=="z"){
          wild <- paste("zip=",gsub(" ","+",secondary),sep="")
        }else{
          wild <- paste("city=",gsub(" ","+",secondary),sep="")
    }
    
    state <- paste("state=",gsub(" ","+",state),sep="") 
    string <-  paste("https://geocoding.geo.census.gov/geocoder/geographies/address?",addy,"&",wild,"&",state,"&benchmark=4&vintage=4&format=json",sep="")
    json_file<-fromJSON(getURL(string))

    #Check if there are results
    if(length(json_file$result$addressMatches)>0){
      
      #If not, kick back an empty dataframe
      if(is.null(json_file$result$addressMatches$coordinates$x[1])==TRUE){
        print("no result")
        return(data.frame(
          address="",
          lat = "",
          lon= "",
          tract = "",
          block = ""))
        
      } else{

        #  Address,lat,lon,tract, block (keep first match)
        address <- as.character(data.frame(json_file$result$addressMatches$matchedAddress)[1,])
        lat <- as.character(json_file$result$addressMatches$coordinates$y[1])
        lon <- as.character(json_file$result$addressMatches$coordinates$x[1])
        tract <- data.frame(json_file$result$addressMatches$geographies$`Census Tracts`)$GEOID[1]
        block <- data.frame(json_file$result$addressMatches$geographies$`2010 Census Blocks`)[1,c("GEOID")]

        return(data.frame(
                address = address,
                lat = lat,
                lon = lon,
                tract = tract,
                block = block))
    
      }
    }
  } #####   end of census_geocoder function


## Geocode the addresses

rpt$match <- NA
rpt$lat <- NA
rpt$lon <- NA
rpt$tract <- NA
rpt$block <- NA

for (i in 1:nrow(rpt)) {
  answer <- NULL
  attempt <- 1
  while( is.null(answer) && attempt <= 3 ) {
    if (attempt>1){print(paste("attempted", attempt))}
    attempt <- attempt + 1
    try(
      answer <- census_geocoder(rpt$Address[i], "z", rpt$Zip[i], "TX")
    )
  } 
  print(paste(i,":",rpt$Address[i], rpt$Zip[i], answer[1]))
  if (!is.null(answer) && nchar(answer$address)>0){
    rpt$match[i] <- answer[[1]]
    rpt$lat[i] <- answer[[2]]
    rpt$lon[i] <- answer[[3]]
    rpt$tract[i] <- answer[[4]]
    rpt$block[i] <- answer[[5]]
  }
  Sys.sleep(1)
}

##  Fix stupid Clutch City thing

rpt$match <- str_replace(rpt$match, "CLUTCH CITY", "HOUSTON")

```


##  Finish up

All that remains is to save the final dataframe. This is all in a self-contained R file that will be run on a cron.

```{r write out weekly}

## Save output

outputfile <- paste(Start_Date,End_Date,sep="_")
outputfile <- str_replace_all(outputfile,"/","_")

print(paste("=========",outputfile,"=========="))

#    Disable this for purposes of this blog post
##saveRDS(rpt, paste("/home/ajackson/Dropbox/Rprojects/CityPermits/",outputfile,".rds", sep=""))

print("------ and we are done")


```

Note that I insert a few messages that will appear in the log file.

##  The cron jobs

I have three cron jobs to take care of the whole process. The first one runs the code above to create and save a data frame with one week's worth of data. That job actually runs rather slowly, due to the geocoding. That is not a fast process. For that reason the next job kicks off some 4 hours later.

```
# m h  dom mon dow   command
 * 16  *   *   2    /usr/lib/R/bin/Rscript '/home/ajackson/Dropbox/Rprojects/CityPermits/ReadPermits.R'  >> '/home/ajackson/Dropbox/Rprojects/CityPermits/ReadPermits.log' 2>&1
 * 20  *   *   2    /usr/lib/R/bin/Rscript '/home/ajackson/Dropbox/Rprojects/CityPermits/merge_permits.R'  >> '/home/ajackson/Dropbox/Rprojects/CityPermits/MergePermits.log' 2>&1
 * 21  *   *   2    /usr/bin/weex mylinux
```

## Merging the files

After waiting 4 hours for the geocoding to happen, then we run the merge step to append all the files together, and more importantly, add the superneighborhood, making use of the appended coordinates. 

```{r merge}

googlecrs <- 4326


inpath <- "/home/ajackson/Dropbox/Rprojects/CityPermits"
path <- "/home/ajackson/mirrors/ajackson/Permits/data"

# Read files in, do date and numeric conversion, and then join together

DF <- list.files(path=inpath, pattern="^20[12].*rds$") %>%
  purrr::set_names(.) %>%
  map_dfr(~ readRDS(file.path(inpath, .))) %>%
  rename(Description=Comments, Date=Permit_Date) %>% 
  mutate(Date=as.Date(Date, format= "%Y/%m/%d"),
         lat=as.numeric(lat),
         lon=as.numeric(lon))

#   fix clutch city

DF$match <- str_replace(DF$match, "CLUTCH CITY", "HOUSTON")

#   filter out rows with no coordinates
DF <- DF %>% filter(!(is.na(DF$lat) | is.na(DF$lon)))

#  Create a temporary sf data frame for doing the intersects
# set longitudes as the first column and latitudes as the second
dat <- data.frame(Longitude=DF$lon, Latitude=DF$lat, match=DF$match, stringsAsFactors = FALSE)

dat <- st_as_sf(dat, coords=c("Longitude", "Latitude"), crs=googlecrs, agr = "identity")

# prep superneighborhoods

temp <- readRDS("/home/ajackson/Dropbox/CrimeStats/SuperNeighborhoodPolys.rds") %>% 
  mutate(SNBNAME=str_to_title(SNBNAME))
#   find points in polygons
#   since superneighborhoods don't overlap, let's just grab the array index
#   instead of creating a huge matrix
a <- st_intersects(dat, temp, sparse = TRUE)

# Replace empty values with 89 - there are some areas not in a neighborhood
a <- unlist(replace(a, !sapply(a, length),89))
# Now add super neighborhood 89 as NA
temp[89,] <- temp[88,]
temp$SNBNAME[89] <- "None"

#   Append the super neighborhood field to the data frame
DF$SuperNeighborhood <- temp$SNBNAME[a]

###   save master file

# commented out for blog
##saveRDS(DF, paste0(path, "/MasterPermits.rds"))


```

## Upload

As the final step, the master file is uploaded to the web by using weex.

The viewer is pretty usable, but still under development. I need to make the links between the table and the map much tighter.

