---
title: Un-winding the addition of prison COVID cases to county totals
author: Alan Jackson
date: '2020-06-15'
slug: un-winding-the-addition-of-prison-covid-cases-to-county-totals.en-us
categories:
  - Prison
  - Data Fitting
tags:
  - Covid-19
  - Texas
keywords:
  - tech
comments: no
showMeta: no
showActions: no
---

## Introduction

Towards the end of May the state of Texas suddenly began adding the number of
COVID-19 cases detected amongst prison inmates to the county totals for the 
counties in which the prisons resided. However, they have not indicated if
they did this change on a single day, or if it may have taken place over
several days, for different prisons. In this bit of work, I will try to ferret
out what they did as best I can, so that I may best correct my own data.

Fortunately I have been collecting the prison data for some time and have a 
fairly complete set of data. I will also play with interpolating that data
to fill in the holes I have, in particular when the URL changed in early
June.

The state also changed their reporting of prison cases at about the same time,
changing what numbers they reported. I will see if I can also reconcile those
changes to generate a single, consistent dataset.

```{r setup, include=FALSE}

library(tidyverse)
library(ggplot2)
library(lubridate)

path <- "/home/ajackson/Dropbox/Rprojects/Covid/"
archive <- "/home/ajackson/Dropbox/Rprojects/Datasets/TexasPrisons/"

Covid_data <- readRDS(paste0(path,"Covid.rds"))

New_prison <- readRDS(paste0(path, "Prisons_2.rds"))
Old_prison <- readRDS(paste0(path, "Prisons.rds"))

prison_loc <- readRDS(paste0(archive, "Prison_Locations.rds"))

prison_loc$Unit_Name <- str_replace(prison_loc$Unit_Name,"Ft. Stockton", "Fort Stockton")
prison_loc$Unit_Name <- str_replace(prison_loc$Unit_Name,"Hospital Galveston", "Hosp Galv")
prison_loc$Unit_Name <- str_replace(prison_loc$Unit_Name,"LeBlanc", "Leblanc")
prison_loc$Unit_Name <- str_replace(prison_loc$Unit_Name,"\\.", "")
prison_loc$Unit_Name <- str_replace(prison_loc$Unit_Name, "Mountain", "Mt.")

Old_prison$Unit <- str_replace(Old_prison$Unit, "Hospital Galveston", "Hosp Galv")
Old_prison$Unit <- str_replace(Old_prison$Unit, "ETTF", "East Texas")
Old_prison$Unit <- str_replace(Old_prison$Unit, "1", "I")
Old_prison$Unit <- str_replace(Old_prison$Unit, "3", "III")
Old_prison$Unit <- str_replace(Old_prison$Unit, "4", "IV")
Old_prison$Unit <- str_replace(Old_prison$Unit, "LeBlanc", "Leblanc")
Old_prison$Unit <- str_replace(Old_prison$Unit, "Lindsay", "Lindsey")
Old_prison$Unit <- str_replace(Old_prison$Unit, "\\.", "")
Old_prison$Unit <- str_replace(Old_prison$Unit, "Sansaba", "San Saba")
Old_prison$Unit <- str_replace(Old_prison$Unit, "Moore", "Moore,")
Old_prison$Unit <- str_replace(Old_prison$Unit, "Mountain", "Mt.")

New_prison <- left_join(New_prison, prison_loc, by=c("Unit"="Unit_Name")) %>% 
  select(-Unit_Code, -Operator, -Gender, -Type, -Region, -City)

bad_units=c("Bambi", "Local Hospital", "No Longer in Custody")
Old_prison <- left_join(Old_prison, prison_loc, by=c("Unit"="Unit_Name")) %>% 
  select(-Unit_Code, -Operator, -Gender, -Type, -Region, -City) %>% 
  filter(!str_detect(Unit, paste(bad_units, collapse="|")))

```

##  Join prison datasets

I have two prison datasets, one prior to May 29, and one starting May 29,
when the website changed and the categories being counted changed. So I need
to combine these into one consistent dataset.

The old set has variables like Pending_Tests, Negative_Tests, Positive_Tests,
and Recovered. The new dataset has Offender Active Cases and 
Offender Recovered. It is not clear if Positive_Tests can count one individual
multiple times or not. I fear that the answer may be yes. Perhaps when we 
combine the datasets it will become clearer.

```{r combine, warning=FALSE, message=FALSE}

Old_prison <- Old_prison %>% 
  mutate(Inmate_cases=Positive_Tests+Recovered) %>% 
  select(Date, Unit, County, Inmate_cases)

New_prison <- New_prison %>% 
  # get total cases
  mutate(Inmate_cases=`Offender Active Cases`+`Offender Recovered`) %>% 
  select(Date, Unit, County, Inmate_cases)

Prison <- rbind(Old_prison, New_prison)


Prison %>% 
  group_by(Unit) %>% 
    mutate(maxi=max(Inmate_cases)) %>% 
  ungroup() %>% 
  filter(maxi>50) %>%
  filter(Date>lubridate::ymd("2020-05-20")) %>% 
  ggplot(aes(x=Date, y=Inmate_cases, color=Unit)) + 
  geom_line() +
  geom_point()


```

Well, I'd say it is still astonishingly unclear. If tests were being counted
multiple times per person, the total should have dropped after May, but for
the most part they have bumped up. I think a lot of that is that they tend
to test people now in large batches, and then all the tests arrive at once
adding a large jump to the numbers. In any event, it appears that Positive
Tests may mean one per inmate. 

##  Cases by county

Let's start by looking at cases by county for both prisons and the general
population, and see how they compare.

```{r cases by county, warning=FALSE, message=FALSE}

#   First sum prison cases within each county

Prison <- Prison %>% 
  # Sum by county
  group_by(Date, County) %>% 
    summarise(Inmate_cases=sum(Inmate_cases)) %>% 
  ungroup() %>% 
  filter(Inmate_cases>0)

#   Check data

Prison %>% 
  filter(Inmate_cases>150) %>% 
  ggplot(aes(x=Date, y=Inmate_cases, color=County)) +
  geom_line() +
  geom_point() +
  labs(title="Inmate Cases by County")

df <- left_join(Prison, Covid_data, by=c("County", "Date"))
  
df %>% 
  group_by(County) %>% 
    mutate(delta=Cases-lag(Cases)) %>% 
    replace_na(list(delta=0)) %>% 
    mutate(size=max(delta)/mean(delta)) %>% 
  ungroup() %>% 
  filter(size>10) %>% 
  ggplot(aes(x=Date, y=delta, color=County)) +
  geom_point() +
  geom_line() +
  geom_line(aes(y=Inmate_cases), linetype="dotted" ) +
  labs(title="Daily change in total cases (delta), with Inmate cases overlain (dotted)")
  
  
df %>% 
  group_by(County) %>% 
    mutate(delta=Cases-Inmate_cases) %>% 
    replace_na(list(delta=0)) %>% 
    mutate(size=max(abs(delta))) %>% 
  ungroup() %>% 
  filter(size<200) %>% 
  ggplot(aes(x=Date, y=delta, color=County)) +
  geom_point() +
  geom_line() +
  geom_line(aes(y=Inmate_cases), linetype="dotted" )+
  labs(title="Daily change in total cases (delta), with Inmate cases overlain (dotted)")

#   Look at top 25 counties to see if a pattern emerges

df <- left_join(Covid_data, Prison, by=c("County", "Date"))
  
df %>% 
  group_by(County) %>% 
    summarise(maxincases=max(Inmate_cases, na.rm=TRUE),
              maxcases=max(Cases, na.rm=TRUE)) %>%
    mutate(fraction=maxincases/(maxincases+maxcases)) %>% 
  ungroup() %>% 
  arrange(desc(fraction)) %>% 
  mutate(rank=row_number()) %>% 
  filter(rank<=25) %>% 
  select(County) -> bigcounties

Bigdata <- df %>% 
  filter(County %in% bigcounties$County)

Bigdata %>% 
  ggplot(aes(x=Date, y=Inmate_cases)) + 
  geom_line() + 
  geom_point() + 
  geom_line(aes(y=Cases), linetype="dashed" ) +
  facet_wrap(~ County, scales = "free_y")
  
```

In the plots above, the points show the inmate numbers, the dashed lines
show total reported cases in the county. In many cases they seem to 
correlate, in many they do not, and in some it is just not clear.

These plots seem to help indicate when and where prison numbers began to be 
added to each county. It appears that I will have to look at each county in 
detail to flag the exact dates to use. And I wonder if those might also
vary within a county by prison unit.

There may also be a day or two lag between test results getting posted
for a prison unit, and those results being propagated to the county.

So let's plot, on a shorter time frame each individual county, and determine
when the prison data was added, and at what lag. And then replot that county
subtracting out the prison data to make sure it looks reasonable.

```{r each county, warning=FALSE, message=FALSE}

for (county in bigcounties$County) {
  p <- df %>% 
    filter(County == county) %>% 
    filter(Date>lubridate::ymd("2020-05-15")) %>% 
    ggplot(aes(x=Date, y=Inmate_cases)) +
    geom_point(aes(color="Inmates"))+
    geom_line(aes(color="Inmates"), linetype="dashed") +
    geom_point(aes(y=Cases, color="Original"))+
    geom_line(aes(y=Cases, color="Original"))+
    scale_color_manual(name="", values=c("red", "black"))+
    labs(title=county, y="Cases")
  
  print(p)
}

Adjusted <- tribble(
  ~County,           ~Date,     ~Lag,
  "Jones",         "2020-05-28", 3,
  "Anderson",      "2020-06-16", 0,
  "Walker",        "2020-04-09", 1,
  "Medina",        "2020-06-06", 0,
  "Rusk",          "2020-05-30", 1,
  "Grimes",        "2020-05-31", 2,
  "Coryell",       "2020-04-26", 2,
  "Terry",         "2020-04-26", 1,
  "Houston",       "2020-04-26", 1,
  "Pecos",         "2020-04-26", 1,
  "Bee",           "2020-04-26", 2,
  "Bowie",         "2020-04-26", 1,
  "Jefferson",     "2020-06-10", 0,
  "Brazoria",      "2020-06-16", 0
)

Adjusted$Date <- lubridate::ymd(Adjusted$Date)

for (county in Adjusted$County) {
  Lag <- Adjusted$Lag[Adjusted$County==county]
  Start_date <- Adjusted$Date[Adjusted$County==county]
  foo <- df %>% #   shift dates by lag
    filter(County == county) %>% 
    mutate(Date=Date+Lag) %>% 
    select(Date, Inmate_cases)
  
  foo[foo$Date<Start_date,]$Inmate_cases <-0 
  foo <- foo %>% mutate(id=row_number())
  #   interpolate gaps
  #foo$Inmate_cases <- zoo::na.approx(foo$Inmate_cases, foo$id, na.rm=FALSE)
  
  
 p <-
    df %>% 
    filter(County == county) %>% 
      select(-Inmate_cases, -Deaths) %>% 
    left_join(., foo, by="Date") %>% 
    mutate(Adjusted=Cases-Inmate_cases) %>% 
    ggplot(aes(x=Date, y=Inmate_cases)) +
    geom_point(aes(color="Inmates"))+
    geom_line(aes(color="Inmates"), linetype="dashed") +
    geom_point(aes(y=Cases, color="Original"))+
    geom_line(aes(y=Cases, color="Original"))+
    geom_point(aes(y=Adjusted, color="Adjusted"))+
    geom_line(aes(y=Adjusted, color="Adjusted"))+
    scale_color_manual(name="", values=c("red", "black", "blue"))+
    labs(title=county, y="Cases")
  
  print(p)
  }


```

Moderate success. Anderson, Rusk, Pecos, Bee, Bowie, Jefferson, and 
Brazoria look pretty good when corrected. Walker, Grimes, Medina, and 
Houston are a bit dodgy. Jones, Coryell, and Terry just don't work. 

So we have a mess. Some counties the prison cases have clearly been added
after a specific date. Some counties they appear to have never been added.
And for some counties it isn't clear what the heck is going on.

But clearly the prison cases, because of delayed batch testing, are 
causing huge jumps in some of the county totals, rendering that data 
nearly useless for most purposes. Subtracting the prison cases does not
appear to be very robust, and given the difficulty in even gathering
the prison data in a reliable fashion, it seems an unwise path to take.

However, I think I can remove the jumps - assuming they are due to prison
reporting. While not perfect, that should improve the data from it's current
state, which is nearly unusable. So let's explore that strategy.

```{r jump removal, warning=FALSE, message=FALSE}

foo <- df %>% 
  group_by(County) %>% 
    mutate(delta=Cases-lag(Cases)) %>% 
    replace_na(list(delta=0)) %>% 
    mutate(size=max(Cases)) %>% 
  mutate(Scaled_delta=delta/size) %>% 
  ungroup()  

foo %>% 
  filter(County %in% bigcounties$County) %>%
  filter(size>50) %>% 
  ggplot(aes(x=delta)) +
  geom_histogram() +
  facet_wrap(~ County, scales = "free_x")
  
#  lag > 25 and size > 50 and delta > 20%

Prison_counties <- c("Jones", "Anderson", "Walker", "Medina", "Rusk",
                     "Grimes", "Coryell", "Houston", "Pecos", "Angelina",
                     "Bowie", "Jefferson", "Brazoria")

#   De-step

foo <- foo %>% 
  arrange(County) %>% 
  filter(County %in% Prison_counties) %>%
  filter(size>50) %>% 
  group_by(County) %>% 
    mutate(Threshold=as.numeric((abs(delta)>25)&(abs(delta)/Cases>0.10))*delta) %>% 
    mutate(Threshold=cumsum(Threshold)) %>%  
  ungroup() %>% 
  mutate(Adjusted=Cases-Threshold) 
  
  for (county in Prison_counties) {
    p <- foo %>% filter(County==county) %>% 
    ggplot(aes(x=Date, y=Inmate_cases)) +
    geom_point(aes(color="Inmates"))+
    geom_line(aes(color="Inmates"), linetype="dashed") +
    geom_point(aes(y=Cases, color="Original"))+
    geom_line(aes(y=Cases, color="Original"))+
    geom_point(aes(y=Adjusted, color="Adjusted"))+
    geom_line(aes(y=Adjusted, color="Adjusted"))+
    scale_color_manual(name="", values=c("red", "black", "blue"))+
    labs(title=county, y="Cases")
  
    print(p)
  }

```

That looks pretty good, especially after I restricted the counties I
apply that logic to. Basically, if a jump is greater than 25 and greater
than 10%, then remove it. The downside is that the numbers now have lost a 
lot of meaning - they represent the county at large with some or most of
the inmate cases removed, but then it isn't totally clear what they
represented before the adjustment. I think this should help make the
time-series behavior more interpretable, even if the absolute numbers
are uncertain. 




