---
title: "Vaccine Reluctance and politics"
author: "Alan Jackson"
date: '2021-05-01'
slug: vaccine-reluctance-and-politics.en-us
categories:
- Covid-19
- Voting
tags:
- Covid-19
- Texas
keywords: tech
---


```{r setup, include=FALSE}

library(tidyverse)
library(ggplot2)
library(lubridate)
library(broom)
library(gt)
library(ggrepel)

#   Directory where data is stored

DataCovid <- "/home/ajackson/Dropbox/Rprojects/Covid/Today_Data/"
DataVotes <- "/home/ajackson/Dropbox/Rprojects/Voting/"
DataDemo <- "/home/ajackson/Dropbox/Rprojects/Curated_Data_Files/Texas_Demographics/"

#   Build Votes by county tibble

Votes <- read_csv(paste0(DataVotes, 
                         "CountybyCountyCanvassReport2020Pres.csv")) %>% 
  rename(Candidate="CANDIDATE NAME", 
         "Office"="OFFICE NAME", 
         "County"="COUNTY NAME",
         Votes="TOTAL VOTES PER OFFICE PER COUNTY") %>% 
  filter(grepl("BIDEN|TRUMP", Candidate)) %>% 
  select(Candidate, County, Votes) %>% 
  pivot_wider(names_from=Candidate, values_from=Votes) %>% 
  rename(Biden=2, Trump=3) %>% 
  mutate(Blueness=Biden/(Biden+Trump))

Covid <- read_rds(paste0(DataCovid, "Today_County_calc.rds"))

out_path <-  "/home/ajackson/Dropbox/Rprojects/Covid/"
Vaccine <- read_rds(paste0(out_path, "Vaccinations.rds")) %>% 
  filter(County!="NULL", 
         County!="Other", 
         County!="*Other", 
         County!="* Other", 
         County!="Texas", 
         !is.na(County),
         County != "Federal Long-Term Care Vaccination Program",
         County != "Federal Pharmacy Retail Vaccination Program") %>% 
  group_by(County) %>% 
    mutate(Pop_teen=max(Pop_teen, na.rm=TRUE)) %>% 
  ungroup()

#   Add total population to file

Vaccine <- Covid %>% 
  select(County, Population) %>% 
  unique() %>% 
  rename(Pop_total=Population) %>% 
  right_join(., Vaccine, by="County")

today_date <- stamp("Aug 1, 2021")(last(Vaccine$Date))

```

## Vaccine Reluctance

Let's look at Texas counties and test various factors for correlations
to the vaccination rate.

We'll primarily look at the rate of the first vaccination, since there are a
variety of reasons why someone might not get the second dose.

Let's start with the raw rates of vaccination by county.

```{r rates}

Vaccine %>% 
  mutate(Pct_one_dose=People_one_dose/Pop_total) %>% 
  ggplot(aes(x=Date, y=Pct_one_dose, color=County)) +
    geom_line(show.legend = FALSE) +
    labs(x="Date", y="First Dose Percentage", title="Texas counties Vaccine Progress")



```
Hmmm...   let's do a little cleanup. We'll pull out the offenders first and look 
at them.

```{r offenders, message=FALSE, warning=FALSE}

foo <- 
Vaccine %>% 
  mutate(Pct_one_dose=People_one_dose/Pop_total) %>% 
  filter(!is.na(Pct_one_dose)) %>% 
  group_by(County) %>% 
    mutate(maxdose=max(Pct_one_dose)) %>% 
  ungroup() %>% 
  select(County, Date, People_one_dose, Pct_one_dose, maxdose) %>% 
  filter(maxdose>0.7)  

#   Cleanup

#   Set specific errors to na

Vaccine$People_one_dose[(Vaccine$County=="Cochran") & 
                      (Vaccine$Date>"2021-02-05")] <- NA
Vaccine$People_one_dose[(Vaccine$County=="Cottle") & 
                      (Vaccine$Date>"2020-12-27") & 
                      Vaccine$Date<"2021-03-24"] <- NA
Vaccine$People_one_dose[(Vaccine$County=="Edwards") & 
                      (Vaccine$Date>"2021-03-13") & 
                      Vaccine$Date<"2021-04-01"] <- NA

#   Delete data showing decrease and interpolate to infill

Vaccine <- 
  Vaccine %>% 
  group_by(County) %>% 
    mutate(Daily_dose=People_one_dose-lag(People_one_dose)) %>% 
  ungroup()

Vaccine$Daily_dose <- replace_na(Vaccine$Daily_dose, 0)

Vaccine$People_one_dose[Vaccine$Daily_dose<0] <- NA

Vaccine <- Vaccine %>% 
    #mutate(Daily_dose=zoo::na.approx(Daily_dose, na.rm=FALSE)) %>% 
    #mutate(People_one_dose=zoo::na.approx(People_one_dose, na.rm=FALSE))  
    mutate(Daily_dose=imputeTS::na_interpolation(Daily_dose, maxgap=5)) %>%   
    mutate(People_one_dose=imputeTS::na_interpolation(People_one_dose, maxgap=5))  
  
#   Replot 

Vaccine %>% 
  mutate(Pct_one_dose=100*People_one_dose/Pop_total) %>% 
  ggplot(aes(x=Date, y=Pct_one_dose, color=County)) +
    geom_line(show.legend = FALSE) +
    labs(x="Date", y="First Dose Percentage", title="Texas counties Vaccine Progress")


```

I think that looks better. At least the nonsensical data is gone.

Now let's look at the distribution.

```{r dist, message=FALSE, warning=FALSE}

Vaccine %>% 
  #filter(Date=="2021-05-01") %>% 
  filter(Date==last(Date)) %>% 
  mutate(Pct_one_dose=100*People_one_dose/Pop_total) %>% 
  #arrange(Pct_one_dose)
  ggplot(aes(x=Pct_one_dose)) +
    geom_histogram() +
    labs(title="Percent Vaccinated With One Dose By Texas County",
         x="Percent Vaccinated",
         subtitle=paste("As of", today_date))

```

Let's look at the progress to getting two doses. We'll plot the number 
who are late getting a second dose by summing first doses up to 3 weeks ago,
and subtracting that from the total fully vaxed.

```{r missed second, message=FALSE, warning=FALSE}

Vaccine %>% 
  mutate(One_dose=if_else((Date<=today()-28), People_one_dose, 0)) %>% 
  group_by(County) %>% 
    summarise(Full=sum(People_fully, na.rm = TRUE), 
              One=sum(One_dose, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(Missed_pct=-100*(Full-One)/One) %>% 
  filter(Missed_pct>0, Missed_pct<100) %>% 
  ggplot(aes(x=Missed_pct)) + 
  geom_histogram() + 
  labs(x="Percent Missing Second Dose",
       y="Number of counties",
       title="Estimated Percent of First Dose Missing Second, by County")

Vaccine %>% 
  mutate(One_dose=if_else((Date<=today()-28), People_one_dose, 0)) %>% 
  group_by(County) %>% 
    summarise(Full=sum(People_fully, na.rm = TRUE), 
              One=sum(One_dose, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(Missed_pct=-100*(Full-One)/One) %>% 
  filter(Missed_pct>15, Missed_pct<100)

```
Let's take a look at the data by regions

```{r regions, message=FALSE, warning=FALSE}

DataArchive <- "/home/ajackson/Dropbox/mirrors/ajackson/SharedData/"
MSA_raw <- readRDS(paste0(DataArchive, "Texas_MSA_Pop_Counties.rds")) 

#  Add vaccine to MSA's 

MSA <- MSA_raw %>% 
  unnest(Counties) %>% 
  rename(County=Counties) %>% 
  left_join(Vaccine, ., by="County") %>% 
  group_by(MSA, Date) %>% 
    summarise(People_one_dose=sum(People_one_dose, na.rm = TRUE),
              People_fully=sum(People_fully, na.rm = TRUE),
              Pop_total=sum(Pop_total, na.rm = TRUE),
              Population=unique(Population)) %>% 
    mutate(Daily_dose=People_fully-lag(People_fully)) %>% 
  ungroup() %>% 
  mutate(Pct_fully=100*People_fully/Pop_total) 

MSA$med <- zoo::rollmedian(MSA$Pct_fully, 9, 
                    fill=c(0, NA, last(MSA$Pct_fully)))  

# Pull out top 5 and bottom 5

MSA_extremes <- 
  MSA %>% 
  group_by(MSA) %>% 
    summarize(med=max(med), MSA=unique(MSA)) %>% 
    arrange(-med) %>% 
  ungroup() 

  
MSA_top <- slice_head(MSA_extremes, n=4)
MSA_bot <- slice_tail(MSA_extremes, n=4)
MSA_topbot <- bind_rows(MSA_top, MSA_bot)

MSA_extremes <- inner_join(MSA, MSA_topbot, by="MSA") %>% 
  rename(med=med.x, ordering=med.y) %>% 
  arrange(-ordering)

#  MSA$Pct_one_dose[MSA$Daily_dose<0] <- NA
# Plot region progress

MSA %>% 
  ggplot(aes(x=Date, y=med, color=MSA)) + 
  geom_line(aes(group=MSA),colour = alpha("grey", 0.7),
            show.legend = FALSE) + 
  geom_line(data=MSA_extremes,
            aes(color=fct_reorder(MSA, -ordering))) + 
  labs(x="Date",
       y="Percent Fully Vaxed",
       title="MSA regions and Vaccination history",
       subtitle=paste("As of", today_date))


```

Let's plot new cases vs. Vaccine status by County

```{r cases vs status, message=FALSE, warning=FALSE}

Vaccine_last <- Vaccine %>% 
  group_by(County) %>% 
    summarise(Vax=last(100*People_fully/Pop_total), 
              County=unique(County)) %>% 
  ungroup()

Cases_last <- Covid %>% 
  group_by(County) %>% 
    summarise(New_cases=last(avg_active_cases_percap), County=unique(County)) %>% 
  ungroup()

Vax_Cases <- left_join(Cases_last, Vaccine_last, by="County")

Vax_Cases %>% 
  ggplot(aes(x=New_cases, y=Vax)) + 
  geom_point() + 
  geom_smooth(method="lm") + 
  scale_x_continuous(trans='log10') + 
  labs(x="Active Cases per 100,000",
       y="Percent Vaccinated",
       title="Active Cases vs. Percent Vaccinated by County in Texas",
       subtitle=paste("As of", today_date))

```

Hmmm.... that doesn't make sense. Slightly more cases where there are more
people vaccinated?

Let's try something different. Let's look at the positive case rate only 
in the unvaccinated cohort.

```{r Cases per unvaxed, message=FALSE, warning=FALSE}

Vaccine_last <- Vaccine %>% 
  group_by(County) %>% 
    summarise(Vax=last(People_fully), 
              Pop=last(Pop_total),
              County=unique(County)) %>% 
  ungroup()

Cases_last <- Covid %>% 
  group_by(County) %>% 
    arrange(Date) %>% 
    summarise(Active=last(active_cases), 
              County=unique(County)) %>% 
  ungroup()

Vax_Cases <- left_join(Cases_last, Vaccine_last, by="County") %>% 
  mutate(New_cases=Active*1.e5/(Pop-Vax), Vax_pct=100*Vax/Pop)  

Vax_Cases %>% 
  ggplot(aes(y=New_cases, x=Vax_pct)) + 
  geom_point() + 
  geom_smooth(method="lm") + 
  scale_y_continuous(trans='log10') + 
  labs(y="Active Cases per 100,000 Un-Vaxxed",
       x="Percent Vaccinated",
       title="Active Cases vs. Percent Vaccinated by County in Texas",
       subtitle=paste("As of", today_date))

```
Still doesn't make sense. Maybe the individual county stats are too noisy -
small rural counties imply small numbers imply noisy data.

Let's combine counties into MSA's to get better behaved statistics.

```{r combine to MSA, message=FALSE, warning=FALSE}

MSA_case <- MSA_raw %>% 
  unnest(Counties) %>% 
  rename(County=Counties) %>% 
  left_join(Covid, ., by="County") %>% 
  group_by(MSA, Date) %>% 
    summarise(Cases=sum(Cases, na.rm = TRUE),
              New_cases=sum(new_cases, na.rm = TRUE),
              Deaths=sum(Deaths, na.rm = TRUE),
              New_Deaths=sum(new_deaths, na.rm = TRUE),
              Active=sum(active_cases, na.rm = TRUE),
              new_tests=sum(new_tests, na.rm=TRUE)
              ) %>% 
  ungroup() 

MSA_final <- 
MSA_case %>% 
  group_by(MSA) %>% 
    slice(tail(row_number(),21)) %>%
    summarize(avg=mean(New_cases, na.rm=TRUE),
              New_Deaths=mean(New_Deaths, na.rm=TRUE),
              new_tests=mean(new_tests, na.rm=TRUE)) %>% 
  ungroup()

MSA_vax <- 
  MSA %>% 
  group_by(MSA) %>% 
    slice(tail(row_number(),21)) %>%
    summarize(Fully=mean(People_fully, na.rm=TRUE),
              Pop=last(Pop_total),
              Tot_Pop=last(Population)) %>% 
  ungroup()
  

MSA_all <- left_join(MSA_final, MSA_vax, by="MSA") %>% 
  mutate(UnvaxCase=avg*1.e5/(Pop-Fully),
         Vax_pct=100*Fully/Pop,
         Cases_per_100k=1.0e5*avg/Tot_Pop,
         Cases_per_unvax=1.0e5*avg/(Tot_Pop-Fully),
         Percap_tests=new_tests*1.0e5/Tot_Pop,
         Percap_deaths_unvax=New_Deaths*1.0e5/(Tot_Pop-Fully)
         ) 

#           Tests vs vax
p1 <- MSA_all %>% 
  #filter(Cases_per_100k<10) %>% 
  ggplot(aes(y=Percap_tests, x=Vax_pct)) + 
  geom_point() + 
  geom_smooth(method="lm") + 
  labs(y="Tests per 100,000",
       x="Percent Vaccinated",
       title="Avg Covid Tests vs. Percent Vaccinated by Metro Area in Texas",
       subtitle=paste("As of", today_date))

#         Cases vs Vaxed
p2 <- MSA_all %>% 
  #filter(Cases_per_100k<10) %>% 
  ggplot(aes(y=Cases_per_100k, x=Vax_pct)) + 
  geom_point() + 
  geom_smooth(method="lm") + 
  labs(y="Avg Cases per 100,000",
       x="Percent Vaccinated",
       title="Avg Covid Cases vs. Percent Vaccinated by Metro Area in Texas",
       subtitle=paste("As of", today_date))

#         Deaths vs Vaxed
p3 <- MSA_all %>% 
  #filter(Cases_per_100k<10) %>% 
  ggplot(aes(y=Percap_deaths_unvax, x=Vax_pct)) + 
  geom_point() + 
  geom_smooth(method="lm") + 
  labs(y="Deaths per 100,000 Unvaxed",
       x="Percent Vaccinated",
       title="Avg Covid Deaths vs. Percent Vaccinated by Metro Area in Texas",
       subtitle=paste("As of", today_date))

#         Deaths vs Tests
p4 <- MSA_all %>% 
  #filter(Cases_per_100k<10) %>% 
  ggplot(aes(y=Percap_deaths_unvax, x=Percap_tests)) + 
  geom_point() + 
  geom_smooth(method="lm") + 
  labs(y="Deaths per 100,000 Unvaxed",
       x="Tests per 100,000",
       title="Avg Covid Deaths vs. Covid Tests by Metro Area in Texas",
       subtitle=paste("As of", today_date))

print(p1)
print(p2)
print(p3)
print(p4)


#grid.arrange(p1,p2,p3,p4)

```
Well, still a bit mysterious. Fewer tests in lower vaxed counties - is that 
test reluctance or fewer people with symptoms?

Deaths also not making sense - though deaths are a lagging indicator, and I'm
not sure how well they tag deaths back to the county of residence. Do the deaths
counted in the county where they were hospitalized and died?

Now let's try to predict the final vaccination rates based on fitting a
logistic function to the percentages.

```{r unvaxed cases, message=FALSE, warning=FALSE}

#---------------------------------------------------    
#------------------- Fit Logistic function ---------
#---------------------------------------------------    
fit_logistic <- function(indep="med", # independent variable
                         df=MSA,
                         MSA=MSA,
                         r=0.24,
                         output=c("fit", "asym"),
                         projection=10){
  
  #print(paste(":::::::  logistic", MSA))
  
  df$Days <- as.integer(df$Date - ymd("2020-12-28"))

  Asym <- max(df$med)*5
  xmid <- max(df$Days)*2
  scal <- 1/r
  my_formula <- as.formula(paste0(indep, " ~ SSlogis(Days, Asym, xmid, scal)"))
  
  #print("----1----")
  
  ## using a selfStart model
    
    logistic_model <- NULL
    try(logistic_model <- nls(my_formula, 
                              data=df)); # does not stop in the case of error
    
    if(is.null(logistic_model)) {
       case_params <<- list(K=NA, 
                            r=NA, 
                            xmid=NA,
                            xmid_se=NA)
      return()
    }
    
  #print("----2----")
  #print(logistic_model)
  coeffs <- coef(logistic_model)
  xmid_sigma <- 2*summary(logistic_model)$parameters[2,2] # 2 sigma
  #print("----3----")
  #print(coeffs)
  #print(summary(logistic_model))
  
  Cases <- predict(logistic_model, data.frame(Days=df$Days))
  if (output=="fit"){ return(Cases) }
  else {return(coeffs)}
} ###############   end of fit_logistic
  
#   Let's try fitting a Logistic to each region as a way to really smooth
# the curves

eraseme <- 
MSA %>% 
  mutate(Days=as.integer(Date - ymd("2020-12-28"))) %>% 
  select(MSA, Days, med, Date) %>% 
  complete(nesting(MSA), Days=seq(min(Days), max(Days), 1L)) %>% 
  nest(-MSA) %>% 
    mutate(fit=map(data, ~ fit_logistic(df=., MSA=MSA, output="fit"))) %>% 
  unnest(data,fit)

eraseme %>% 
  ggplot(aes(x=Date, color=MSA)) + 
  geom_line(aes(group=MSA, y=fit),colour = alpha("grey", 0.7),
            show.legend = FALSE) + 
  geom_point(aes(y=med)) + 
  labs(x="Date",
       y="Percent Vaxed",
       title="MSA regions and Vaccination history",
       subtitle=paste("As of", today_date))


asymptotes <- 
MSA %>% 
  mutate(Days=as.integer(Date - ymd("2020-12-28"))) %>% 
  select(MSA, Days, med, Date) %>% 
  complete(nesting(MSA), Days=seq(min(Days), max(Days), 1L)) %>% 
  nest(-MSA) %>% 
    mutate(Coeffs = map(data, ~ fit_logistic(df=.,
                                                MSA=MSA, 
                                                output="asym"))) %>% 
  unnest_wider(Coeffs)

asymptotes <- asymptotes %>% 
  select(MSA, Asym, xmid, scal) %>% 
  left_join(., MSA, by="MSA") %>% 
  select(MSA, Asym, xmid, scal, Population, Pop_total) %>% 
  group_by(MSA) %>% 
    filter(row_number()==n()) %>% 
  ungroup() %>% 
  mutate(midDate=xmid + ymd("2020-12-28"))

asymptotes %>% 
  arrange(-Asym) %>% 
  select(MSA, Asym, Pop_total) %>% 
  mutate(Asym=Asym/100) %>% 
  gt() %>% 
  tab_header(title="Texas Regions: Logistic Fit Asymptotes",
             subtitle="Expected final percent vaccinated") %>% 
  fmt_percent(columns=2, decimals=1) %>% 
  fmt_number(columns=3, decimals=0) %>% 
  cols_label(MSA= md("**Metro Stat Area**"),
             Asym= md("**Asymptote**"),
             Pop_total= md("**Eligible (12+) Population**"))
  
asymptotes %>% 
  arrange(-Asym) %>% 
  select(MSA, Asym, Pop_total) %>% 
  mutate(Asym=Asym/100) %>%
  mutate(vaxed=Asym*Pop_total) %>% 
  summarize(allvaxed=sum(vaxed), allpop=sum(Pop_total))


```
Rather depressing. Good looking fits, implying that unless something changes, 
many counties will get no where near herd immunity levels. It looks like a
surge of Delta cases waiting to happen.

##   Let's look at vaccine vs vote

```{r votes, message=FALSE, warning=FALSE}

Votes <- Votes %>% 
  mutate(County=str_to_title(County))

Votes[Votes$County=='Mcculloch',]$County <- "McCulloch"
Votes[Votes$County=='Mclennan',]$County <- "McLennan"
Votes[Votes$County=='Mcmullen',]$County <- "McMullen"

Vax_vote <- left_join(Votes, Vaccine_last, by="County") %>% 
  mutate(Vax_pct=Vax/Pop) 

rects <- data.frame(ymin = -Inf, 
                    ymax = Inf,
                    xmin = c(-Inf,0.45,0.55),  
                    xmax = c(0.45,0.55,Inf),
                    fill = c("red", "magenta", "blue"))
Vax_vote %>% 
  ggplot(aes(y=Vax_pct, x=Blueness)) +
  geom_rect(data = rects, aes(xmin = xmin, 
                              xmax = xmax, 
                              ymin = ymin, 
                              ymax = ymax, 
                              fill = fill), 
            # Control the shading opacity here.
            inherit.aes = FALSE, alpha = 0.15) +
  scale_fill_identity() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  geom_point() +
  geom_smooth(method="lm") +
  geom_text_repel(data=subset(Vax_vote, Vax_pct < 0.30 | 
                          Vax_pct > 0.60 |
                          Blueness > 0.55),
            aes(Blueness,Vax_pct,label=County), hjust=0,
                nudge_x=.008) +
  labs(x="Fraction of Vote For Biden",
       y="Percent Vaccinated",
       title="Vaccination Rate per Texas County",
       subtitle=paste("As of", today_date),
       #subtitle=stamp("March 1, 1999")(today()-1),
       caption="Data from https://dshs.texas.gov/coronavirus/AdditionalData.aspx")
#  geom_text_repel(aes(label=County))
 
#   Vaccination rates vs demographics

Demographics <- readRDS(paste0(DataDemo, "County_Age_Sex.rds"))

Demo_data <- Demographics %>% 
  mutate(Pct_over_65=0.01*(Total_pct_90plus +
                     Total_pct_85to89 +
                     Total_pct_80to84 +
                     Total_pct_75to79 +
                     Total_pct_70to74 +
                     Total_pct_65to69)) %>% 
  select(CNTY_NM, FIPS, Pct_over_65)

Vax_vote <- Vax_vote %>% 
  left_join(., Demo_data, by=c("County"="CNTY_NM")) 
  
Vax_vote %>% 
  ggplot(aes(y=Vax_pct, x=Pct_over_65)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  geom_point() +
  geom_smooth(method="lm") +
  geom_text_repel(data=subset(Vax_vote, Vax_pct < 0.30 | 
                          Vax_pct > 0.60 |
                          Blueness > 0.55),
            aes(Pct_over_65,Vax_pct,label=County), hjust=0,
                nudge_x=.008) +
  labs(x="Percent of Population over 65",
       y="Percent Vaccinated",
       title="Vaccination Rate per Texas County",
       subtitle=paste("As of", today_date),
       #subtitle=stamp("March 1, 1999")(today()-1),
       caption="Data from https://dshs.texas.gov/coronavirus/AdditionalData.aspx")
```

Best correlation I've gotten yet. And the saddest. Politicians encouraging
people to make bad decisions and die for political gain. 

Let's look at the history of case counts vs. politics in 2-month windows.

```{r 2 month windows}

#   Shrink data to what will be needed and sum in 2 month window

df <- 
Covid %>% 
  select(County, Cases, Deaths, Date, Population) %>% 
  mutate(End_date=lubridate::ceiling_date(Date, unit="month")-1) %>% 
  filter(Date==End_date) %>% # pull out last day of every month
  group_by(County) %>% 
    mutate(Monthly_cases=pmax((Cases-lag(Cases, default=0)),0),
           Monthly_deaths=pmax((Deaths-lag(Deaths, default=0)),0)
           ) %>% 
  ungroup() %>% 
  mutate(bucket = case_when(
    month(Date) < 3  ~ "Jan-Feb",
    month(Date) < 5  ~ "Mar-Apr",
    month(Date) < 7  ~ "May-Jun",
    month(Date) < 9  ~ "Jul-Aug",
    month(Date) < 11 ~ "Sep-Oct",
    month(Date) < 13 ~ "Nov-Dec",
    TRUE             ~ "Unknown"
  )) %>% 
  mutate(bucket=ifelse(year(Date)<2021, 
                       paste(bucket, "2020"), 
                       paste(bucket, "2021"))) %>% 
  group_by(County,bucket) %>% 
    summarize(Cases=sum(Monthly_cases),
              Deaths=sum(Monthly_deaths),
              Population=unique(Population)) %>% 
  ungroup()

df$bucket <- factor(df$bucket, levels=c("Jan-Feb 2020", "Mar-Apr 2020",
                                        "May-Jun 2020", "Jul-Aug 2020",
                                        "Sep-Oct 2020", "Nov-Dec 2020",
                                        "Jan-Feb 2021", "Mar-Apr 2021",
                                        "May-Jun 2021", "Jul-Aug 2021",
                                        "Sep-Oct 2021", "Nov-Dec 2021",
                                        "Unknown 2021"))
#   Add in votes

df <- left_join(df, Votes, by="County")

#  plots

df %>%
  mutate(Cases_per_cap=1.e5*Cases/Population) %>% 
  ggplot(aes(x=Blueness, y=Cases_per_cap)) + 
  geom_point() + 
  geom_smooth(method="lm") +
  scale_y_continuous(trans='log10') + 
  facet_wrap(vars(bucket), scales="free_y") + 
  labs(x="Fraction In County Voting for Biden",
       y="Cases per 100,000",
       title="Covid Cases by Texas County in Two Month Intervals",
       subtitle=paste("As of", today_date))
  

df %>%
  mutate(Cases_per_cap=1.e5*Deaths/Population) %>% 
  ggplot(aes(x=Blueness, y=Cases_per_cap)) + 
  geom_point() + 
  geom_smooth(method="lm") +
  scale_y_continuous(trans='log10') + 
  facet_wrap(vars(bucket), scales="free_y") + 
  labs(x="Fraction In County Voting for Biden",
       y="Deaths per 100,000",
       title="Covid Deaths by Texas County in Two Month Intervals",
       subtitle=paste("As of", today_date))
  
#    What is case vs death rates for various counties?  
  
df %>%
  mutate(Cases_per_cap=Deaths/Cases) %>% 
  ggplot(aes(x=Blueness, y=Cases_per_cap)) + 
  geom_point() + 
  geom_smooth(method="lm") +
  scale_y_continuous(trans='log10') + 
  facet_wrap(vars(bucket), scales="free_y") + 
  labs(x="Fraction In County Voting for Biden",
       y="Deaths per Case",
       title="Covid Deaths per Case by Texas County in Two Month Intervals",
       subtitle=paste("As of", today_date))
  
  

```
It does look like supporting Trump has amounted to a death sentence for many.

Let's bring in the hospitalization data and add that into the mix

```{r hospital}

#   First Trauma Service Areas since that is how the hospital data is set

inpath <- "/home/ajackson/Dropbox/Rprojects/Curated_Data_Files/Texas_Trauma_Service_Areas/"

TSA <- readRDS(paste0(inpath, "Trauma_Service_Areas.rds"))

#   Read in hospitalization data

inpath <- "/home/ajackson/Dropbox/Rprojects/CovidTempData/DailyBackups/"

Hospital <- readRDS(paste0(inpath,"2022-03-21_Hospital.rds"))

Last_date <- last(Hospital$Date)

#   Attach TSA to Vaccine and Case files, then summarize by TSA

Vax_TSA <- Vaccine %>% 
  select(County, Date, People_fully, Pop_teen, Pop_total) %>% 
  left_join(., TSA, by="County") %>% 
  rename(TSA_Area = Area_name) %>% 
  group_by(TSA_Area, Date) %>% 
    summarise(People_fully=sum(People_fully, na.rm=TRUE),
              Pop_teen=sum(Pop_teen, na.rm = TRUE),
              Pop_total=sum(Pop_total, na.rm=TRUE)) %>% 
  left_join(., Hospital, by=c("TSA_Area", "Date")) %>% 
  select(-TSA_ID) %>% 
  filter(Date<=Last_date)

#   Now add in votes

Votes[Votes$County=='Dewitt',]$County <- "DeWitt"

Votes_TSA <- Votes %>% 
  left_join(., TSA, by="County") %>% 
  rename(TSA_Area = Area_name) %>% 
  group_by(TSA_Area) %>% 
    summarize(Biden=sum(Biden, na.rm=TRUE),
              Trump=sum(Trump, na.rm=TRUE)) %>% 
  mutate(Blueness=Biden/(Trump+Biden))

Vax_TSA <- Vax_TSA %>% 
  left_join(., Votes_TSA, by="TSA_Area")

#   Most recent values

Vax_TSA_now <- Vax_TSA %>% 
  mutate(Hospitalized=as.numeric(Hospitalized)) %>% 
  filter(Date==Last_date) %>% 
  mutate(Vax_pct=People_fully/Pop_total) %>%
  mutate(Hosp_percap=1.e5*Hospitalized/Pop_total)


#   Now plot like crazy

#         Vax rate vs. vote
Vax_TSA_now %>% 
  ggplot(aes(y=Vax_pct, x=Blueness)) +
  geom_rect(data = rects, aes(xmin = xmin, 
                              xmax = xmax, 
                              ymin = ymin, 
                              ymax = ymax, 
                              fill = fill), 
            # Control the shading opacity here.
            inherit.aes = FALSE, alpha = 0.15) +
  scale_fill_identity() +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  geom_point() +
  geom_smooth(method="lm") +
  geom_text_repel(data=subset(Vax_TSA_now, Vax_pct < 0.38 | 
                          Vax_pct > 0.65 |
                          Blueness > 0.55),
            aes(Blueness,Vax_pct,label=TSA_Area), hjust=0,
                nudge_x=.005) +
  labs(x="Fraction of Vote For Biden",
       y="Percent Vaccinated",
       title="Vaccination Rate per Texas Trauma Service Area",
       subtitle=paste("As of", today_date))
      # subtitle=stamp("March 1, 1999")(Last_date))  

#         Hospitalization rate vs. vote

Vax_TSA_now %>% 
  ggplot(aes(y=Hosp_percap, x=Blueness)) +
  geom_rect(data = rects, aes(xmin = xmin, 
                              xmax = xmax, 
                              ymin = ymin, 
                              ymax = ymax, 
                              fill = fill), 
            # Control the shading opacity here.
            inherit.aes = FALSE, alpha = 0.15) +
  scale_fill_identity() +
  #scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
  geom_point() +
  geom_smooth(method="lm") +
  geom_text_repel(data=subset(Vax_TSA_now, Hosp_percap > 10),
            aes(Blueness,Hosp_percap,label=TSA_Area), hjust=0,
                nudge_x=.005) +
  labs(x="Fraction of Vote For Biden",
       y="Covid Hospitalizations per 100,000",
       title="Covid Hospitalization Rate per Texas Trauma Service Area",
       subtitle=paste("As of", today_date))
       #subtitle=stamp("March 1, 1999")(Last_date))  

#         Hospitalization rate vs. Vaccination Rate

Vax_TSA_now %>% 
  ggplot(aes(y=Hosp_percap, x=Vax_pct)) +
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
  geom_point() +
  geom_smooth(method="lm") +
  geom_text_repel(data=subset(Vax_TSA_now, Hosp_percap > 10),
            aes(Vax_pct,Hosp_percap,label=TSA_Area), hjust=0,
                nudge_x=.005) +
  labs(x="Percent Vaccinated",
       y="Covid Hospitalizations per 100,000",
       title="Covid Hospitalization Rate per Texas Trauma Service Area",
       subtitle=paste("As of", today_date))
       #subtitle=stamp("March 1, 1999")(Last_date))  

#   Plots vs time

Vax_TSA <- Vax_TSA %>% 
  mutate(Hospitalized=as.numeric(Hospitalized)) %>% 
  mutate(Vax_pct=People_fully/Pop_total) %>%
  mutate(Hosp_percap=1.e5*Hospitalized/Pop_total)

# Pull out top 5 and bottom 5

Vax_TSA_extremes <- 
  Vax_TSA %>% 
  group_by(TSA_Area) %>% 
    summarize(Hosp_percap=last(Hosp_percap), TSA_Area=unique(TSA_Area)) %>% 
    arrange(-Hosp_percap) %>% 
  ungroup() 

  
TSA_top <- slice_head(Vax_TSA_extremes, n=4)
TSA_bot <- slice_tail(Vax_TSA_extremes, n=4)
TSA_topbot <- bind_rows(TSA_top, TSA_bot)

TSA_extremes <- inner_join(Vax_TSA, TSA_topbot, by="TSA_Area") %>% 
  rename(Hosp_percap=Hosp_percap.x, ordering=Hosp_percap.y) %>% 
  arrange(-ordering)

# Plot region progress

Vax_TSA %>% 
  ggplot(aes(x=Date, y=Hosp_percap, color=TSA_Area)) + 
  geom_line(aes(group=TSA_Area),colour = alpha("grey", 0.7),
            show.legend = FALSE) + 
  geom_line(data=TSA_extremes,
            aes(color=fct_reorder(TSA_Area, -ordering))) + 
  labs(x="Date",
       y="Covid Hospitalizations per 100,000",
       title="Covid Hospitalization Rate per Texas Trauma Service Area",
       subtitle=paste("As of", today_date))
       #subtitle=stamp("March 1, 1999")(Last_date))  

#   Vaccination pct vs time by TSA


# Pull out top 5 and bottom 5 by vax

Vax_TSA_extremes <- 
  Vax_TSA %>% 
  group_by(TSA_Area) %>% 
    summarize(Vax_pct=last(Vax_pct), TSA_Area=unique(TSA_Area)) %>% 
    arrange(-Vax_pct) %>% 
  ungroup() 

  
TSA_top <- slice_head(Vax_TSA_extremes, n=4)
TSA_bot <- slice_tail(Vax_TSA_extremes, n=4)
TSA_topbot <- bind_rows(TSA_top, TSA_bot)

TSA_extremes <- inner_join(Vax_TSA, TSA_topbot, by="TSA_Area") %>% 
  rename(Vax_pct=Vax_pct.x, ordering=Vax_pct.y) %>% 
  arrange(-ordering)

# Plot region progress

Vax_TSA %>% 
  ggplot(aes(x=Date, y=Vax_pct, color=TSA_Area)) + 
  geom_line(aes(group=TSA_Area),colour = alpha("grey", 0.7),
            show.legend = FALSE) + 
  geom_line(data=TSA_extremes,
            aes(color=fct_reorder(TSA_Area, -ordering))) + 
  geom_hline(yintercept=0.50, color="black", size=0.5, 
             show.legend = FALSE) +
  labs(x="Date",
       y="Vaccination percentage",
       title="Vaccination Rate Rate per Texas Trauma Service Area",
       subtitle=paste("As of", today_date))
       #subtitle=stamp("March 1, 1999")(Last_date))  



```

Clearer picture in these areas. Fewer vaccinations imply more hospitalizations.

Votes for Trump imply more hospitalizations.

Next let's try out some trajectory plots.

```{r trajectory}


# Pull out top 5 and bottom 5 by hospitalization

Vax_TSA_extremes_hosp <- 
  Vax_TSA %>% 
  group_by(TSA_Area) %>% 
    summarize(Hosp_percap=last(Hosp_percap), TSA_Area=unique(TSA_Area)) %>% 
    arrange(-Hosp_percap) %>% 
  ungroup() 

  
TSA_top_hosp <- slice_head(Vax_TSA_extremes_hosp, n=4)
TSA_bot_hosp <- slice_tail(Vax_TSA_extremes_hosp, n=4)
TSA_topbot_hosp <- bind_rows(TSA_top_hosp, TSA_bot_hosp)

TSA_extremes_hosp <- inner_join(Vax_TSA, TSA_topbot_hosp, by="TSA_Area") %>% 
  rename(Hosp_percap=Hosp_percap.x, ordering=Hosp_percap.y) %>% 
  arrange(-ordering)

Vax_TSA %>% 
  ggplot(aes(x=Vax_pct, 
             y=Hosp_percap,
             color=TSA_Area)) +
  geom_line(aes(group=TSA_Area),colour = alpha("grey", 0.7),
            show.legend = FALSE) + 
  geom_line(data=TSA_extremes_hosp,
            aes(color=fct_reorder(TSA_Area, -ordering))) + 
  geom_vline(xintercept=0.50, color="black", size=0.5, 
             show.legend = FALSE) +
  labs(x="Percent Vaccinated (12+ Population)",
       y="Covid Hospitalizations per 100,000",
       title="Hospitalization/Vaccination Trajectories per Texas Trauma Service Area",
       subtitle=paste("As of", today_date))


```



```{r cases tsa}

#   Attach TSA to Case file, then summarize by TSA

Case_TSA <- Covid %>% 
  select(County, Date, Deaths, Cases, new_cases, new_deaths, Population) %>% 
  left_join(., TSA, by="County") %>% 
  rename(TSA_Area = Area_name) %>% 
  filter(Date>lubridate::ymd("2021-01-01")) %>% 
  group_by(TSA_Area, Date) %>% 
    summarise(Deaths=sum(Deaths, na.rm=TRUE),
              Cases=sum(Cases, na.rm = TRUE),
              Pop=sum(Population, na.rm = TRUE),
              new_cases=sum(new_cases, na.rm = TRUE),
              new_deaths=sum(new_deaths, na.rm = TRUE)
              ) %>% 
  mutate(cases_percap=Cases/Pop*100000) %>% 
  mutate(deaths_percap=(Deaths-first(Deaths))/Pop*100000) %>% 
  mutate(deaths_percase=Deaths/Cases) %>% 
  left_join(., Vax_TSA, by=c("TSA_Area", "Date")) %>% 
  filter(!is.na(Vax_pct)) %>% 
  mutate(Vax_pct=Vax_pct*100) %>% 
  group_by(TSA_Area) %>% 
     mutate(Vax_pct=zoo::rollmedian(Vax_pct,
                                          5,
                                          fill=c(0, NA, last(Vax_pct)))) %>% 
  ungroup() %>% 
  filter(Date<=Last_date)


# Pull out top 5 and bottom 5 by hospitalization

Vax_TSA_extremes_case <- 
  Case_TSA %>% 
  group_by(TSA_Area) %>% 
    summarize(deaths_percap=last(deaths_percap), TSA_Area=unique(TSA_Area)) %>% 
    arrange(-deaths_percap) %>% 
  ungroup() 

  
TSA_top_case <- slice_head(Vax_TSA_extremes_case, n=4)
TSA_bot_case <- slice_tail(Vax_TSA_extremes_case, n=4)
TSA_topbot_case <- bind_rows(TSA_top_case, TSA_bot_case)

TSA_extremes_case <- inner_join(Case_TSA, TSA_topbot_case, by="TSA_Area") %>% 
  rename(deaths_percap=deaths_percap.x, ordering=deaths_percap.y) %>% 
  arrange(-ordering)

Case_TSA %>% 
  filter(Date>lubridate::ymd("2021-01-01")) %>% 
#  group_by(TSA_Area) %>% 
#     mutate(deaths_percap=deaths_percap-first(deaths_percap)) %>% 
#  ungroup() %>% 
  ggplot(aes(x=Vax_pct, 
             y=deaths_percap,
             color=TSA_Area)) +
  geom_line(aes(group=TSA_Area),colour = alpha("grey", 0.7),
            show.legend = FALSE) + 
  geom_line(data=TSA_extremes_case,
            aes(color=fct_reorder(TSA_Area, -ordering))) + 
 # geom_vline(xintercept=0.50, color="black", size=0.5, 
 #            show.legend = FALSE) +
  labs(x="Vaccination Percent",
       y="Covid Deaths per 100,000",
       title="Vaccination/Death Trajectories per Texas Trauma Service Area Since January",
       subtitle=paste("As of", today_date))



```
Vacination rates before and after full approval on August 23.

```{r approval}

#  Let's look at 4 week tranches before and after approval

Approval = ymd("2021-08-13")

foo <- 
Vax_TSA %>% 
  mutate(Approval_week=(ceiling(as.numeric(Date-Approval)/7))) %>% 
  group_by(TSA_Area, Approval_week) %>% 
    summarize(All_dose=last(People_fully)-first(People_fully),
              Pop=first(Pop_total),
              n=n()) %>% 
  filter(All_dose>0) %>% 
  mutate(Week_date=Approval+Approval_week*n) %>% 
  mutate(All_dose_percap=All_dose/(Pop/100000)) %>% 
  filter(n==7)  
 
# Pull out top 5 and bottom 5 by last Vax

Vax_TSA_extremes_case <- 
  foo %>% 
  group_by(TSA_Area) %>% 
    summarize(All_dose_percap=last(All_dose_percap),
              TSA_Area=unique(TSA_Area)) %>% 
    arrange(-All_dose_percap) 

  
TSA_top_case <- slice_head(Vax_TSA_extremes_case, n=4)
TSA_bot_case <- slice_tail(Vax_TSA_extremes_case, n=4)
TSA_topbot_case <- bind_rows(TSA_top_case, TSA_bot_case)

TSA_extremes_case <- inner_join(foo, TSA_topbot_case,
                                by="TSA_Area") %>% 
  rename(All_dose_percap=All_dose_percap.x, 
         ordering=All_dose_percap.y) %>% 
  arrange(-ordering)

 
foo %>% 
  ggplot(aes(y=All_dose_percap, x=Week_date)) +
  #geom_line(aes(color=TSA_Area)) +
  geom_line(aes(group=TSA_Area),colour = alpha("grey", 0.7),
            show.legend = FALSE) + 
  geom_line(data=TSA_extremes_case,
            aes(color=fct_reorder(TSA_Area, -ordering))) + 
  scale_colour_discrete(name="TSA Area") +
  geom_vline(xintercept=Approval, color="black", size=0.5, 
             show.legend = FALSE) +
  geom_text(x=Approval, y=0, label="Full FDA Approval") +
  labs(x="Date",
       y="Number of vaccinations given per 100,000",
       title="Weekly Vaccination in Texas by Trauma Service Area",
       subtitle=paste("As of", today_date))
  
  




```




