---
title: 'Houston Geocoding - Part 2'
author: Alan Jackson
date: '2022-05-19'
slug: [Houston_Geocoding-Part-2]
categories:
  - Geocoding
tags:
  - Houston
  - Mapping
keywords:
  - tech
---

```{r setup, include=FALSE}

library(tidyverse)
library(ggplot2)
library(lubridate)
library(leaflet)
library(GeocodeHou)

googlecrs <- 4326

path <- "/home/ajackson/Dropbox/Rprojects/Curated_Data_Files/CoH_Address_Points/"

knitr::opts_chunk$set(echo = TRUE)
```

## Geocoding Part 2

Let's take the address for the Art Car Museum and use that as our example
address. The first address is correct, the next 5 have a flaw in one of the
fields.

```{r initialize}

#   Address for the Art Car Museum

Test_data <- tribble(
  ~ID, ~Street_num, ~Prefix, ~Street_name, ~Street_type, ~Zipcode,
  "1",       "140",   ""   ,    "Heights",       "BLVD",  "77007",
  "2",       "138",   ""   ,    "Heights",       "BLVD",  "77007",
  "3",       "140",  "W"   ,    "Heights",       "BLVD",  "77007",
  "4",       "140",   ""   ,    "Hieghts",       "BLVD",  "77007",
  "5",       "140",   ""   ,    "Heights",         "LN",  "77007",
  "6",       "140",   ""   ,    "Heights",       "BLVD",  "77070"
)



```

###  Exact Matches

The basic expected way to run the code is to first find all exact matches, and
then use the additional tools to try to repair any failures that occurred.

As can be seen, the correct address has a good match, and the Lat/Long are 
returned. For the next five addresses, a flag noting at which stage the
match failed is returned.

The matching is fairly conservative. I assume that only one field is in error,
and so, for example, for most repairs will restrict the attempt to the given 
zip code, assuming that it is correct. 

For the exact match, the search order for matching is:

1. Zip code
2. Street name
3. Street type
4. Street number
5. Direction prefix

The potential matches are subsetted with each test, and if the match fails
at any stage, that is the failure that is returned. So, for example, the last
address which has the wrong zip code fails with a Street Name failure, because
there is no street name in that wrong zip code that matches, but the zip code
itself is valid.

```{r exact matches}

for (i in 1:nrow(Test_data)){ # first look for exact matches
  tmp <- match_exactly(Test_data[i,]$Street_num, 
                       Test_data[i,]$Prefix,
                       Test_data[i,]$Street_name,
                       Test_data[i,]$Street_type, 
                       Test_data[i,]$Zipcode)
  if (tmp$Success){ #   success
    print(paste("Success", tmp$Lat, tmp$Lon))
  } else { #  Fail exact match
    print(paste("Failed", tmp$Fail))
  }
}

```

### Zip Code

Zip code repair entails looking for an exact address match in any other
zip code. Given that all the other repair routines assume a correct zip code, 
this one should be run first.

The fail message for the correct address simply means that the full address
was found in the supposedly incorrect zip code - normally you would not
try to repair an address you already know is correct.

The incorrect zip code is corrected and the lat/long is returned. 

```{r zipcode}

for (i in 1:nrow(Test_data)){ # first look for exact matches
  tmp <- repair_zipcode(Test_data[i,]$Street_num, 
                        Test_data[i,]$Prefix,
                        Test_data[i,]$Street_name,
                        Test_data[i,]$Street_type, 
                        Test_data[i,]$Zipcode)
  if (tmp$Success){ #   success
    print(paste("Success", tmp$Lat, tmp$Lon, tmp$New_zipcode))
  } else { #  Fail exact match
    print(paste("Failed", tmp$Fail))
  }
}

```

### Street Name

The distance to all names in the given zip code is calculated using the 
generalized Levenshtein (edit) distance. Basically if one letter has to be 
changed to get a match, that is a distance of one. Two letters, a distance of
two. The distance can be specified, but the default of two is pretty safe.

Certain street names are ignored, as they are too prone to error:

- Names of < 4 letters
- Names like "A STREET"
- Names like "12TH"
- County roads (CR 123)
- Farm Roads (FM 123)

If more than one match is found, that is also a failure case.

```{r name}

for (i in 1:nrow(Test_data)){ # first look for exact matches
  tmp <- repair_name(Test_data[i,]$Street_num, 
                        Test_data[i,]$Prefix,
                        Test_data[i,]$Street_name,
                        Test_data[i,]$Street_type, 
                        Test_data[i,]$Zipcode)
  if (tmp$Success){ #   success
    print(paste("Success", tmp$Lat, tmp$Lon, tmp$New_name))
  } else { #  Fail exact match
    print(paste("Failed", tmp$Fail))
  }
}

```

### Street Type

Again, a search will occur within the designated zip code, and all the street
types for the given name and prefix will be compiled. If only a single type is
found, then an exact match is attempted. If that succeeds then the new address
and lat/long are returned.

The street types Court and Circle (CT, CIR) are ignored, as they are too risky.


```{r type}

for (i in 1:nrow(Test_data)){ # first look for exact matches
  tmp <- repair_type(Test_data[i,]$Street_num, 
                        Test_data[i,]$Prefix,
                        Test_data[i,]$Street_name,
                        Test_data[i,]$Street_type, 
                        Test_data[i,]$Zipcode)
  if (tmp$Success){ #   success
    print(paste("Success", tmp$Lat, tmp$Lon, tmp$New_type))
  } else { #  Fail exact match
    print(paste("Failed", tmp$Fail))
  }
}

```

### Street Number

If the street number is not found, then we try to interpolate from nearby
addresses. Which sounds simple, but is a bit tricky. What constitutes
nearby? And given that the locations are building or plat centered, how
should we interpolate?

First we look for address numbers on the same block as the target address,
and on the same side of the street (even or odd). If we find less than 2
addresses, we open up the search to 3 blocks. If there are still less than 2 
numbers, we give up. It could be argued that at this point we could ignore 
the even odd distinction to open up more possibilities, which would work
in some cases, but at the cost of introducing significant bias in the 
interpolated result. More complex interpolation schemes can also be
imagined, but it is unclear how much benefit would be gained. Google uses
some of these to poor effect.

Next we calculate delta lat/longs for pairs of addresses taken in order to
come up with an interpolator. We then use the address nearest to the target
to calculate the new lat/long. Finally we check the distance from the nearest 
point to the target. If it exceeds a threshold (set by a parameter) then the
match is rejected. The default value is 250 feet.


```{r num}

for (i in 1:nrow(Test_data)){ # first look for exact matches
  tmp <- repair_number(Test_data[i,]$Street_num, 
                        Test_data[i,]$Prefix,
                        Test_data[i,]$Street_name,
                        Test_data[i,]$Street_type, 
                        Test_data[i,]$Zipcode)
  if (tmp$Success){ #   success
    print(paste("Success", tmp$Lat, tmp$Lon, tmp$Result))
  } else { #  Fail exact match
    print(paste("Failed", tmp$Fail))
  }
}

```

### Directional Prefix

Similar to Street Type, we will search for the possible prefixes in the given 
zip code. If only one is found, then we declare success.

In the example, we are unsuccessful. Why is that? Because within the zip code
77007, there is one, maybe two blocks of Heights Blvd south of Washington
Ave where it becomes S. Heights Blvd, before becoming Waugh. So there is a 
fundamentally unresolvable ambiguity if we know the prefix is wrong.

```{r prefix}

for (i in 1:nrow(Test_data)){ # first look for exact matches
  tmp <- repair_prefix(Test_data[i,]$Street_num, 
                        Test_data[i,]$Prefix,
                        Test_data[i,]$Street_name,
                        Test_data[i,]$Street_type, 
                        Test_data[i,]$Zipcode)
  if (tmp$Success){ #   success
    print(paste("Success", tmp$Lat, tmp$Lon, tmp$New_prefix))
  } else { #  Fail exact match
    print(paste("Failed", tmp$Fail))
  }
}

```

### Summary

I have developed an R package for geocoding within the city of Houston. In tests
it has performed well. I feel confident that when it gives a good result, that 
answer is quite trustworthy - much more so than Google or the Census bureau.

In my tests against the city permit file, out of about 130,000 addresses, about
122,000 were matched exactly. Of the roughly 8000 failed matches, I was able
to repair and recover roughly half of those addresses. Of the remaining
failures, maybe 10% were not even addresses, so impossible to recover even 
manually. Most of the rest had multiple issues that would be difficult to
sort out.



